---
title: "Create figure showing the evolution of big teams over time and the topics they study"
author: "Anonymized for peer review (NC)"
reviewer: "Anonymized for peer review (JT)"
output: word_document
editor_options: 
  chunk_output_type: console
---

This code creates a figure illustrating (a) frequency of big team psychology efforts over time, and (b) the topics they study.

# Setup
```{r setup, include= F}
# setup
source("setup.R")

# specify data directory
data_dir <- Sys.getenv("data_dir")

# specify directory
i_am('btpsych24a_timetopic.Rmd')
```

# Open time data
```{r open.alex, include = F}
# open OpenAlex data
DF <-
  readRDS(file.path(data_dir, "openalex", "btpsych24a_data_bib_processed.Rds"))
```

## Create percentile labels
For the plot, we want labels to mention both (a) the numerical range of the authorship category, and (b) what this category means in terms of percentile
```{r perc.labels, include = F}
# calculate percentiles
per.DF <- DF %>%
  select(authors.n) %>%
  mutate(perc = percent_rank(authors.n) * 100)

# create column for storing bin labels
DF$bin.label <- NA

# create label for size.bin == 1 - 4
DF[DF$size.bin == "1 - 4", ]$bin.label <-
  paste0(
    per.DF %>%
      filter(authors.n >= 1, authors.n < 5) %>%
      slice_min(perc) %>%
      max %>%
      round(0),
    " - ",
    per.DF %>%
      filter(authors.n >= 1, authors.n < 5) %>%
      slice_max(perc) %>%
      max %>%
      round(0),
    " percentile"
  )

# create label for size.bin == 5 - 9
DF[DF$size.bin == "5 - 9", ]$bin.label <-
  paste0(
    per.DF %>%
      filter(authors.n >= 5, authors.n < 10) %>%
      slice_min(perc) %>%
      max %>%
      round(0),
    " - ",
    per.DF %>%
      filter(authors.n >= 5, authors.n < 10) %>%
      slice_max(perc) %>%
      max %>%
      round(1),
    " percentile"
  )

# create label for size.bin == 10 - 19
DF[DF$size.bin == "10 - 19", ]$bin.label <-
  paste0(
    per.DF %>%
      filter(authors.n >= 10, authors.n < 20) %>%
      slice_min(perc) %>%
      max %>%
      round(1),
    " - ",
    per.DF %>%
      filter(authors.n >= 10, authors.n < 20) %>%
      slice_max(perc) %>%
      max %>%
      round(2),
    " percentile"
  )

# create label for size.bin == 20 - 29
DF[DF$size.bin == "20 - 29", ]$bin.label <-
  paste0(
    per.DF %>%
      filter(authors.n >= 20, authors.n < 30) %>%
      slice_min(perc) %>%
      max %>%
      round(2),
    " - ",
    per.DF %>%
      filter(authors.n >= 20, authors.n < 30) %>%
      slice_max(perc) %>%
      max %>%
      round(2),
    " percentile"
  )

# create label for size.bin == 30 - 39
DF[DF$size.bin == "30 - 39", ]$bin.label <-
  paste0(
    per.DF %>%
      filter(authors.n >= 30, authors.n < 40) %>%
      slice_min(perc) %>%
      max %>%
      round(3),
    " - ",
    per.DF %>%
      filter(authors.n >= 30, authors.n < 40) %>%
      slice_max(perc) %>%
      max %>%
      round(3),
    " percentile"
  )

# create label for size.bin == 40 - 49
DF[DF$size.bin == "40 - 49", ]$bin.label <-
  paste0(
    per.DF %>%
      filter(authors.n >= 40, authors.n < 50) %>%
      slice_min(perc) %>%
      max %>%
      round(3),
    " - ",
    per.DF %>%
      filter(authors.n >= 40, authors.n < 50) %>%
      slice_max(perc) %>%
      max %>%
      round(3),
    " percentile"
  )

# create label for size.bin == 50 - 59
DF[DF$size.bin == "50 - 59", ]$bin.label <-
  paste0(
    per.DF %>%
      filter(authors.n >= 50, authors.n < 60) %>%
      slice_min(perc) %>%
      max %>%
      round(3),
    " - ",
    per.DF %>%
      filter(authors.n >= 50, authors.n < 60) %>%
      slice_max(perc) %>%
      max %>%
      round(3),
    " percentile"
  )

# create label for size.bin == 60 - 69
DF[DF$size.bin == "60 - 69", ]$bin.label <-
  paste0(
    per.DF %>%
      filter(authors.n >= 60, authors.n < 70) %>%
      slice_min(perc) %>%
      max %>%
      round(3),
    " - ",
    per.DF %>%
      filter(authors.n >= 60, authors.n < 70) %>%
      slice_max(perc) %>%
      max %>%
      round(3),
    " percentile"
  )

# create label for size.bin == 70 - 79
DF[DF$size.bin == "70 - 79", ]$bin.label <-
  paste0(
    per.DF %>%
      filter(authors.n >= 70, authors.n < 80) %>%
      slice_min(perc) %>%
      max %>%
      round(3),
    " - ",
    per.DF %>%
      filter(authors.n >= 70, authors.n < 80) %>%
      slice_max(perc) %>%
      max %>%
      round(3),
    " percentile"
  )

# create label for size.bin == 80 - 89
DF[DF$size.bin == "80 - 89", ]$bin.label <-
  paste0(
    per.DF %>%
      filter(authors.n >= 80, authors.n < 90) %>%
      slice_min(perc) %>%
      max %>%
      round(3),
    " - ",
    per.DF %>%
      filter(authors.n >= 80, authors.n < 90) %>%
      slice_max(perc) %>%
      max %>%
      round(3),
    " percentile"
  )

# create label for size.bin == 90 - 99
DF[DF$size.bin == "90 - 99", ]$bin.label <-
  paste0(
    per.DF %>%
      filter(authors.n >= 90, authors.n < 100) %>%
      slice_min(perc) %>%
      max %>%
      round(3),
    " - ",
    per.DF %>%
      filter(authors.n >= 90, authors.n < 100) %>%
      slice_max(perc) %>%
      max %>%
      round(3),
    " percentile"
  )

# create label for size.bin == 100 +
DF[DF$size.bin == "100 +", ]$bin.label <-
  paste0(
    " < ",
    per.DF %>%
      filter(authors.n >= 100) %>%
      slice_min(perc) %>%
      max %>%
      round(3),
    " percentile"
  )

# combine percentile and bin size information
DF <- DF %>%
  mutate(bin.label = paste0(size.bin, " co-authors", "\n", bin.label))

# delete vestigial
rm(per.DF)
```

## Calculate percent growth
```{r}
tmp <- DF %>%
  # focus on growth between 2011 and 2020
  filter(publication_year > 2010, publication_year < 2021) %>%

  # distinguish between first and second half of that decade
  mutate(pub.period = if_else(publication_year > 2015, "later", "earlier")) %>%

  # calculate change over time for each size.bin
  group_by(size.bin, pub.period) %>%
  summarise(n = n()) %>%
  pivot_wider(
    names_from = 'pub.period',
    values_from = n,
    id_cols = 'size.bin'
  ) %>%
  rowwise() %>%
  mutate(
    per.growth = (later - earlier) / earlier,
    per.growth = per.growth * 100,
    per.growth = round(per.growth, 0)
  )
```

## Create facet labels
```{r}
# create facet labels
f.label <- unique(DF$bin.label)
names(f.label) <- unique(DF$size.bin)

# reorder labels (move "5 - 9" to second position and "100 +" to end)  #JT: adjusted text
time.DF <- DF %>%
  mutate(bin.label = factor(bin.label)) %>%
  mutate(
    bin.label = fct_relevel(
      bin.label,
      str_subset(levels(bin.label), "5 - 9 co-authors"),
      after = 1
    ),
    bin.label = fct_relevel(
      bin.label,
      str_subset(levels(bin.label), "100 +"),
      after = Inf
    )
  )

# calculate n's for facet labels
n.DF <- time.DF %>%
  group_by(bin.label) %>%
  summarise(size.bin = unique(size.bin), n = n())

# add proportion for for facet labels
n.DF <- left_join(n.DF, tmp, by = "size.bin")

# delete vestigial
rm(tmp)
```

## Identify where to place annotations
Growth annotations
```{r}
# filter data
label.data <-
  time.DF %>%
  filter(bin.label == str_subset(levels(bin.label), "5 - 9 co-authors"))

# estimate density
label.pos <-
  density(label.data$publication_year, bw = 3)

# turn density into counts
label.pos$y = label.pos$y * nrow(label.data)

# identify position of max point of curve
label.coord.growth <-
  data.frame(
    bin.label = unique(label.data$bin.label),
    x.early = 2011.059,
    y.early = approx(x = label.pos$x, y = label.pos$y, xout = 2011.059)$y,
    x.mid = 2015.070,
    y.mid = approx(x = label.pos$x, y = label.pos$y, xout = 2015.070)$y,
    x.later = 2020.045,
    y.later = approx(x = label.pos$x, y = label.pos$y, xout = 2020.045)$y
  )

# delete vestigial
rm(label.data, label.pos)
```

Indexing lag annotations
```{r}
# filter data
label.data <-
  time.DF %>%
  filter(bin.label == str_subset(levels(bin.label), "1 - 4 co-authors"))
# create density plot
label.pos <-
  density(label.data$publication_year, bw = 3)

# turn density into counts
label.pos$y = label.pos$y * nrow(label.data)

# identify position of max point of curve
label.coord.lag <-
  data.frame(
    bin.label = unique(label.data$bin.label),
    x.pos = label.pos$x[which.max(label.pos$y)],
    y.pos = label.pos$y[which.max(label.pos$y)]
  )

# delete vestigial
rm(label.data, label.pos)
```

## Create time figure
```{r time.fig}
fig.tim <-
  ggplot(data = time.DF, aes(x = publication_year, y = ..count..)) +

  # facet by size.bin
  facet_wrap(~bin.label, scales = "free_y") +

  # overall density
  geom_line(
    aes(color = bin.label),
    stat = "density",
    bw = 3,
    linewidth = 1.25,
    alpha = 0.4
  ) +

  # add category sample size labels
  geom_text(
    data = n.DF,
    aes(x = 2000.5, y = Inf, label = paste0("N = ", n)),
    hjust = 0,
    vjust = 1.5,
    size = 2.75
  ) +

  # add growth labels
  geom_text(
    data = n.DF,
    aes(x = 2000.5, y = Inf, label = paste0(per.growth, "% growth")),
    hjust = 0,
    vjust = 4,
    size = 2.75
  ) +

  # add percent growth annotation
  ## earliest in 10 year mark
  geom_segment(
    data = label.coord.growth,
    aes(x = x.early, xend = x.early, y = y.early, yend = -Inf),
    linetype = 'dotted',
    linewidth = .5,
    color = "grey50"
  ) +

  ## middle
  geom_segment(
    data = label.coord.growth,
    aes(x = x.mid, xend = x.mid, y = y.mid, yend = -Inf),
    linetype = 'dotted',
    linewidth = .25,
    color = "grey50"
  ) +

  ## end of 10 year mark
  geom_segment(
    data = label.coord.growth,
    aes(x = x.later, xend = x.later, y = y.later, yend = -Inf),
    linetype = 'dotted',
    linewidth = .5,
    color = "grey50"
  ) +

  ## annotation
  geom_text(
    data = label.coord.growth,
    aes(x = 2015, y = y.mid - 10000),
    label = '10-year\ngrowth',
    color = "grey50",
    size = 2.5
  ) +

  # add index lag annotation
  geom_segment(
    data = label.coord.lag,
    aes(x = x.pos, xend = x.pos, y = y.pos, yend = -Inf),
    linetype = 'dotted',
    linewidth = .5,
    color = "grey50"
  ) +

  geom_segment(
    data = label.coord.lag,
    aes(x = x.pos, xend = x.pos + 5, y = y.pos - 60000, yend = y.pos - 60000),
    linetype = "dotted",
    linewidth = .5,
    arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
    color = "grey50"
  ) +

  geom_text(
    data = label.coord.lag,
    aes(x = x.pos, y = y.pos - 50000),
    label = 'indexing\nlag',
    color = "grey50",
    size = 2.5
  ) +

  # improve aesthetics
  ylab("Count") +
  xlab("Publication Year") +
  theme(
    legend.position = "none",
    legend.title = element_blank(),
    strip.text = element_text(size = 8)
  ) +
  guides(color = guide_legend(override.aes = list(size = 7))) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 3)) +
  scale_x_continuous(n.breaks = 3)

```

# Topics data
Open data
```{r}
# open
DF.topics <- read.csv(file.path(data_dir, 'topic', 'btpsych24a_topic.csv'))

# fix corrupted size bin variable (Excel thought it was a date)
DF.topics <- DF.topics %>%
  mutate(
    size.bin = case_when(
      size.bin == "4-Jan" ~ "1 - 4",
      size.bin == "9-May" ~ "5 - 9",
      size.bin == "19-Oct" ~ "10 - 19",
      TRUE ~ size.bin # Keep original value if no condition is met
    )
  )

# remove incompletes
DF.topics <-
  DF.topics %>%
  filter(topic != "N/A", topic != " N/A", topic != "N//A")

# recode topic
DF.topics <- DF.topics %>%
  mutate(
    # focus on the first two numbers (highest order of the classification scheme)
    topic.trunc = substr(topic, 1, 2),

    # add name to classification scheme
    # see https://www.apa.org/pubs/databases/training/class-codes#:~:text=Every%20record%20in%20APA%20PsycInfo,retrieval%20to%20a%20specific%20category
    topic.trunc = recode(
      topic.trunc,
      "21" = 'General Psychology',
      "22" = 'Psychometrics & Statistics & Methodology',
      "23" = 'Human Experimental Psychology',
      "24" = 'Animal Experimental & Comparative Psychology',
      "25" = 'Physiological Psychology & Neuroscience',
      "26" = 'Psychology & the Humanities',
      "27" = 'Communication Systems',
      "28" = 'Developmental Psychology',
      "29" = 'Social Processes & Social Issues',
      "30" = 'Social Psychology',
      "31" = 'Personality Psychology',
      "32" = 'Psychological & Physical Disorders',
      "33" = 'Health & Mental Health Treatment & Prevention',
      "34" = 'Health & Mental Health Personnel Issues',
      "35" = 'Educational & School Psychology',
      "36" = 'Organizational Psychology & Human Resources',
      "37" = 'Sport Psychology & Leisure',
      "38" = 'Military Psychology',
      "39" = 'Consumer Psychology',
      "40" = 'Engineering & Environmental Psychology',
      "41" = 'Cognitive Psychology & Intelligent Systems',
      "42" = 'Forensic Psychology & Legal Issues'
    )
  )
```

## Create topic figure
```{r}
topic.fig.DF <-
  DF.topics %>%

  # refactor size.bin
  mutate(
    size.bin = factor(size.bin),
    size.bin = fct_relevel(size.bin, "5 - 9", after = 1),
    size.bin = fct_relevel(size.bin, "100 +", after = Inf)
  ) %>%

  # calculate summary statistics (n's and proportions)
  group_by(topic.trunc, size.bin) %>%

  # sort by frequency
  mutate(
    topic.trunc = factor(
      topic.trunc,
      levels = fct_infreq(.$topic.trunc) %>%
        levels %>%
        rev
    )
  ) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(
    n = if_else(
      size.bin == '1 - 4' |
        size.bin == '5 - 9',
      true = 0,
      false = n
    )
  ) %>%
  mutate(prop = n / sum(n))

# calculate standard errors
topic.fig.se <- topic.fig.DF %>%
  group_by(topic.trunc) %>%
  summarise(total.prop = sum(prop), total.n = sum(n)) %>%
  mutate(se = sqrt((total.prop * (1 - total.prop)) / sum(total.n)))

topic.fig <-
  ggplot(data = topic.fig.DF, aes(y = topic.trunc, x = prop)) +

  # bar chart
  geom_bar(aes(fill = size.bin), stat = 'identity', alpha = .4) +

  # error bars
  geom_errorbar(
    data = topic.fig.se,
    aes(x = total.prop, xmin = total.prop - se, xmax = total.prop + se),
    width = .5,
    colour = 'grey50'
  ) +

  # improve aesthetics
  xlab('Proportion of Big Team Records') +
  ylab('Topic') +
  theme(
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = 'none'
  )
```

# Combine plots
```{r}
plot_grid(
  fig.tim,
  topic.fig,
  nrow = 2,
  ncol = 1,
  rel_heights = c(.65, .35),
  labels = c("A", "B")
)
```
