---
title: "Create contributorship figure"
author: "Anonymized for peer review (NC)"
reviewer: "Anonymized for peer review (JT)"
output: word_document
editor_options: 
  chunk_output_type: console
---

# Setup
```{r setup, include= F}
# load libaries
library('tidyverse')
library('here')
library('ggtext')
library('cowplot')

# set plotting theme
theme_set(theme_classic()) 

# specify data directory
data_dir <- Sys.getenv("data_dir")

# turn off scientific notation
options(scipen = 999)
```

# Open OpenAlex data
```{r open.alex, include = F}
DF.oa <- 
  # open data
  readRDS(file.path(data_dir, 
                    "openalex",
                    "btpsych24a_data_bib_processed.Rds")) %>% 
  
  # select variables of interest
  select(id, authors.n, size.bin, citation.p.y)
```

# Open and process contributorship data
Keep in mind that there are some instances where you couldn't actually code
```{r}
DF.c <-
  # open data
  read.csv(file.path(data_dir, 
                'contributorship',
                'btpsych24a_cont_coding.csv'),
           na.string = "N/A") %>% 
  
  # filter
  filter(
    # remove instances without contributorship statements
    !is.na(contributorship.statement),
    contributorship.statement != "",
    
    # remove instances where coders recommended removal 
    remove.revise != 1,
    
    # remove instances where there wasn't anything to code
    !(is.na(conception) &
        is.na(execution) &
        is.na(analysis) &
        is.na(draft.writing) &
        is.na(review.writing) &
        is.na(funding)
      )
    )  %>% 
  
  # select relevant variables
  select(id, contributorship.statement : funding)
```

# Merge OpenAlex and contributorship data
```{r}
DF <- 
  inner_join(DF.c,  #JT: changed to inner join to analyze with partial data set
             DF.oa, 
             by = "id")

# rm(DF.c, DF.oa)
```

# Identify number of contributors per category
```{r}
DF <- DF %>% 
  # pivot longer for easier processing
  pivot_longer(cols = c(conception : funding),
               names_to = 'category',
               values_to = 'statements') %>% 
  
  # count number of authors who have contributed to each category
  mutate(
    n = 
      # if all authors, use full author count
      if_else(statements == "All authors",
              authors.n,
              
              # otherwise, if N/A, leave blank
              if_else(is.na(statements),
                      NA,
                      
                      # otherwise, count number of commas in the extracted contributorship statement and add 1  
                      str_count(statements,
                                  ",") + 1)
                )
    ) %>% 
  
  # calculate proportion of authors who contributed to each category
  mutate(prop.n = n / authors.n) %>% 
  
  # re-arrange category ordering
  mutate(category = factor(category,
                           levels = c('conception',
                                      'execution',
                                      'analysis',
                                      'draft.writing',
                                      'review.writing',
                                      'funding')
                           )
         ) %>% 
  
  # manually fix proportions that exceed 1
  mutate(prop.n = if_else(prop.n > 1,
                          1,
                          prop.n)) %>% 
  
  # remove categories not being analyzed
  filter(category != 'funding', 
         category != 'review.writing')

# identify dominant contribution category
tmp <- 
  DF %>% 
  
  # process each paper separately
  group_by(id) %>%
  
  # grab top two categories
  slice_max(prop.n,
            n = 2,
            with_ties = F) %>% 
  select(id, category, prop.n) %>% 
  
  # identify which is max vs. second max
  mutate(position = row_number()) %>% 
  
  # If the percentage of authors who contributed to the first category is at least 10% more than the percentage of the second, list it as dominant
  pivot_wider(names_from = 'position',
              values_from = 'prop.n') %>% 
  mutate(dominant = 
           if_else(max(`1`, na.rm = T) > max(`2`, na.rm = T) + .1,
                   'yes',
                   'no')) %>% 
  filter(!is.na(`1`)) %>% 
  rename('category.main' = category) %>% 
  select(id, category.main, dominant)

# get descriptive statistics for dominant categories
tmp2 <- tmp %>% 
  filter(dominant == "yes") %>% 
  group_by(category.main) %>% 
  summarise(category.main.prop = n() / (DF$id %>% unique %>% length), # divided by number of unique records
            category.main.prop = round(category.main.prop, 2) * 100)

# merge data with its descriptives
DF.2 <- 
  # merge dataframe (DF) and category descriptives (tmp)
  left_join(DF, 
            tmp,
            by = "id") %>% 
  
  # merge dominant category descriptives (tmp2)
  left_join(.,
            tmp2,
            by = "category.main") %>% 
  
  # update labels
  mutate(category.main = paste0(category.main, 
                                " dominant (",
                                category.main.prop,
                                "%)")) %>% 
  mutate(category.main = factor(category.main,
                                levels = unique(category.main), 
                                labels = unique(category.main)
                           )
         ) %>% 
  
  # remove unnecessary variables
  select(-citation.p.y)

# export processed data
saveRDS(DF.2,
        file.path(data_dir, 
             'contributorship',
             'btpsych24a_cont_processed.Rds'))
```

# Create figure
## Plot the percentage of contributing authors within each category
Prep data
```{r}
overall.fig.DF <-  
  DF.2 %>% 
  group_by(category) %>% 
  summarise(M = mean(prop.n,
                     na.rm = T),
            sd = sd(prop.n,
                    na.rm = T),
            n = length(unique(id)),
            se = sd / sqrt(n)) %>% 
  rename('prop.n' = 'M')
```

Plot
```{r}
overall.fig <- 
  ggplot(overall.fig.DF,
         aes(x = category,
             y = prop.n)) +
  
  # violin and box plots
  geom_violin(data = DF.2,
              aes(fill = category),
              alpha = .5,
              color = NA,
              width = 1.2) +
  
  geom_boxplot(data = DF.2,
               width= 0.1, 
               color= "grey70", 
               alpha= 0.5,
               fill = NA,
               outlier.shape = NA) +
  
  # improve aesthetics
  theme(legend.position = "none") +
  labs(x = "", 
       y = "proportion of authors") +
  scale_x_discrete(labels = c('conceptualization',
                              'execution',
                              'analysis',
                              'writing')) +
  scale_fill_brewer(palette = "Pastel2") +
  scale_color_brewer(palette = "Pastel2")
```

## Plot the contribution patterns of articles with a dominant contribution category
Prep data
```{r}
# focus on instances where a collaboration was focused on a specific type of contribution
#JT: suggestion: select cases where a specific contribution category predominated in number of author contributions 
sub.fig.DF <- DF.2 %>% 
  filter(dominant == 'yes')
```

Plot
```{r}
sub.fig <-
  ggplot(sub.fig.DF,
       aes(x = category,
           y = prop.n)
       ) +
  
  # wrap by category
  facet_wrap(vars(category.main)) +
  
  # create line graphs
  geom_line(aes(group = id,
                color = category.main),
            alpha = .2,
            linewidth = 1) +
  
  # improve aesthetics
  theme(legend.position = "none") +
  labs(x = "type of contribution",
       y = "proportion of authors") +
  scale_x_discrete(labels = c('conceptualization',
                              'execution',
                              'analysis',
                              'writing')) +
  scale_fill_brewer(palette = "Pastel2") +
  scale_color_brewer(palette = "Pastel2")
```

## Combine figures
```{r}
combined.fig <- 
  plot_grid(overall.fig,
            sub.fig,
            nrow = 2,
            rel_heights = c(4, 6),
            labels = c("A", "B"))
```
