---
title: "Process OpenAlex data and prepare other essential data files"
author: "Anonymized for peer review (NC)"
reviewer: "Anonymized for peer review (JT)"
output: word_document
editor_options: 
  chunk_output_type: console
---
This code (a) lightly processes OpenAlex records, (b) exports DOI's to be used in Altmetric data collection, and (c) exports records for contributorship coding.

# Setup
```{r include= F}
# specify data directory
data_dir <- Sys.getenv("data_dir")

# source setup file
source("setup.R")
```

# Open and process OpenAlex data
```{r open.alex, include = F}
# open raw data
DF <-
  readRDS(
    file.path(
      data_dir,
      "openalex",
      "btpsych24a_data_bib_raw.Rds"
    )
  ) %>%
  # filter data
  filter(
    # only include articles
    type == "article",

    # remove records where we don't have author information
    !is.na(author)
  ) %>%
  # identify number of authors and citations per year
  rowwise() %>%
  mutate(
    authors.n = nrow(author),

    # categorize authors counts
    size.bin = cut(authors.n,
      breaks = c(
        -Inf,
        4, 9, 19,
        29, 39,
        49, 59,
        69, 79,
        89, 99,
        Inf
      ),
      labels = c(
        "1 - 4",
        "5 - 9",
        "10 - 19",
        "20 - 29",
        "30 - 39",
        "40 - 49",
        "50 - 59",
        "60 - 69",
        "70 - 79",
        "80 - 89",
        "90 - 99",
        "100 +"
      )
    ),

    # calculate citations per year
    citation.p.y = cited_by_count / (2025 - publication_year)
  ) %>%
  ungroup()
```

# Prep Altmetric DOIs
The below code was used to export DOIs, which were used by JT to export Altmetric files in batches of 25,000 
```{r altmetric.prep, include = F, eval = F}
# bind id information
ids <-
  bind_rows(DF$ids)

# prepare data for export
doi <-
  ids %>%
  # remove cases with no DOI
  filter(!is.na(doi)) %>%
  # select relevant variables
  select(openalex, doi) %>%
  # add tracking variable placeholder
  mutate(submitted = "")

# export data
## first million records
write.csv(doi[1:1000000, ],
  file.path(
    data_dir,
    "altmetric",
    "btpsych24a_bib_doi1.csv"
  ),
  row.names = F
)

## rest of records
write.csv(doi[1000001:1936245, ],
  file.path(
    data_dir,
    "altmetric",
    "btpsych24a_bib_doi2.csv"
  ),
  row.names = F
)
```

# Prep contributorship coding sheet
```{r eval = F}
DF %>%
  # isolate largest team efforts
  filter(
    size.bin != "1 - 9",
    size.bin != "10 - 19"
  ) %>%
  # randomly sample 1000 records
  slice_sample(n = 1000) %>%
  # select relevant variables
  select(id, doi, title, ab) %>%
  # create rows that coders may want
  mutate(
    remove.revise = "",
    contributorship.statement = ""
  ) # %>%

# export (commented out to not override)

# write.csv(file.path(data_dir,'contributorship',
#                'btpsych24a_cont_coding.csv'),
#           row.names = F)
```

# Filter out duplicate records
Duplicate records were discovered on October 18, 2024
```{r}
# pull out records with NA's, as these will be erroneously labeled as 'duplicates'
DF.na <-
  DF %>%
  filter(is.na(doi))

# remove duplicates
DF <-
  DF %>%
  distinct(doi,
    .keep_all = T
  )

# merge dataframes
DF <-
  rbind(DF, DF.na) %>%
  arrange(id)

# delete vestigial
rm(DF.na)
```

# Correct author counts for records with > 100 authors
The OpenAlex API labels these papers as only having 100 authors.
```{r}
# create file if it doesn't already exist (it should)
if (!file.exists(file.path(
  data_dir,
  "openalex",
  "btpsych24a_author.n.correction.csv"
))
) {
  DF %>%
    select(id, authors.n) %>%
    filter(authors.n == 100) %>%
    rename("authors.n.orig" = authors.n) %>%
    mutate(authors.n.fixed = "") %>%
    write.csv(.,
      file.path(
        data_dir,
        "openalex",
        "btpsych24a_author.n.correction.csv"
      ),
      row.names = F
    )
}

# load file
DF.author.correction <-
  read.csv(file.path(
    data_dir,
    "openalex",
    "btpsych24a_author.n.correction.csv"
  )) %>%
  select(-authors.n.orig)

# join
DF <-
  # left join
  left_join(
    x = DF,
    y = DF.author.correction,
    by = "id"
  ) %>%
  # if record's author.n needs to be fixed, fix it
  mutate(
    authors.n =
      if_else(!is.na(authors.n.fixed),
        authors.n.fixed,
        authors.n
      )
  ) %>%
  # delete vestigial variable
  select(-authors.n.fixed)

rm(DF.author.correction)
```

# Use title keywords to filter out frequently ineligible records
```{r}
DF <-
  DF %>%
  filter(
    !str_detect(
      title,
      regex("Transactions of|Transactions on|Conference|Annual meeting|Symposium|Proceedings of|Table of Contents",
        ignore_case = T
      )
    )
  )
```

# Remove duplicated entries identified using a MinHashing technique (August 7th)
```{r}
# read remove list generated by the btpsych24a_dedup.R script
remove_list <- readRDS(file.path(data_dir, "deduplication", "remove_list.Rds"))

# Remove dataframe entries that match items in remove_list
DF <- DF %>%
  filter(!(id %in% remove_list))
```

# Examine other filters that Reviewer 2 recommended (July 25)
```{r}
tmp <- DF %>%
  filter(
    str_detect(
      so,
      regex("arxiv",
        ignore_case = T
      )
    )
  )

DF %>%
  filter(
    str_detect(
      so,
      regex("SSRN",
        ignore_case = T
      )
    )
  ) %>%
  nrow()
```

# Select relevant variables (to reduce size) and identify source's h index

```{r}
# Select variables of interest
DF <-
  DF %>%
  select(
    id, title, author,
    language, doi, publication_year,
    cited_by_count,
    authors.n:citation.p.y, so_id
  )

# read source ID and h-index collected through OpenAlex`s API
OA_sources <-
  readRDS(file.path(
    data_dir,
    "openalex",
    "OA_sources.Rds"
  )) %>%
  rename(so_id = id) %>%
  select(so_id, h_index)

# Join each work and their journal`s h-index by the source's ID
DF <- left_join(DF, OA_sources, by = "so_id")
```

# Export processed OpenAlex data
```{r}
# save
saveRDS(
  DF,
  file.path(
    data_dir,
    "openalex",
    "btpsych24a_data_bib_processed.Rds"
  )
)

# optionally re-load
# DF <-
#     readRDS(file.path(data_dir,
#                       'openalex',
#                       'btpsych24a_data_bib_processed.Rds'))
```


# Export list of big team efforts for topic coding
```{r eval = F}
DF.bts <- DF %>%
  # select random sample
  group_by(size.bin) %>%
  slice_sample(n = 300) %>%
  # create placeholder for manual extraction of topic from APA PsychInfo
  mutate(topic = "") %>%
  # isolate doi information
  mutate(
    doi =
      str_replace(
        string = doi,
        pattern = "https://doi.org/",
        replacement = ""
      )
  ) %>%
  # select relevant variables
  select(id, doi, title, size.bin, topic) %>%
  # arrange
  arrange(desc(size.bin), doi)

# export
DF.bts %>%
  write.csv(
    file.path(
      data_dir,
      "topic",
      "btpsych24a_topiccoding.csv"
    ),
    row.names = F
  )
```
