---
title: "btpsych24a paper refs"
author: "Anonymized for peer review (NC)"
reviewer: "Anonymized for peer review (JT)"
output: word_document
editor_options: 
  chunk_output_type: console
---
This code compiles the numbers quoted in the paper.

# Setup
```{r setup, include= F}
# setup
source("setup.R")

# specify directory
i_am("btpsych24a_paper.refs.Rmd")

# specify data directory
data_dir <- Sys.getenv("data_dir")

# turn off scientific notation
options(scipen = 999)
```

# New (unsorted)
Examine whether mentions of terms varies
```{r}
DF %>%
  mutate(rep = str_detect(string = title, pattern = 'replication')) %>%
  filter(rep == T) %>%
  nrow()

DF %>%
  mutate(rep = str_detect(string = title, pattern = 'meta-analysis')) %>%
  filter(rep == T) %>%
  nrow()
```

```{r}

## Examine whether topic distribution varies
``
`{r}
m.0 <-
  multinom(topic.trunc ~ 1, data = DF.topics)

m.1 <-
  multinom(topic.trunc ~ size.bin, data = DF.topics)

anova(m.0, m.1)

topic.dist <-
  table(DF.topics$size.bin, DF.topics$topic.trunc) %>%
  as.data.frame() %>%
  pivot_wider(names_from = 'Var2', values_from = 'Freq')

rm(m.0, m.1)
```

```

```{r}
DF.oa.filter[DF.oa.filter$unique_institutions == 0, ]$unique_institutions <- NA
DF.oa.filter[DF.oa.filter$unique_country_codes == 0, ]$unique_country_codes <- NA

# plot(DF.oa.filter$authors.n, DF.oa.filter$unique_institutions)
# plot(DF.oa.filter$authors.n, DF.oa.filter$unique_country_codes)

DF.oa.filter %>%
  group_by(size.bin) %>%
  summarise(
    mean.i =
      mean(unique_institutions,
        na.rm = T
      ),
    median.i =
      median(unique_institutions,
        na.rm = T
      ),
    mean.c =
      mean(unique_country_codes,
        na.rm = T
      ),
    median.c =
      median(unique_country_codes,
        na.rm = T
      )
  )

# calculate correlations
cor.test(
  DF.oa.filter$authors.n,
  DF.oa.filter$unique_institutions,
  method = "spearman"
)

cor.test(
  DF.oa.filter$authors.n,
  DF.oa.filter$unique_country_codes,
  method = "spearman"
)

# quick thought: is team size, country, or institution a better predictor?
cor.test(
  DF.oa.filter$authors.n,
  DF.oa.filter$citation.p.y,
  method = "spearman"
)

cor.test(
  DF.oa.filter$unique_institutions,
  DF.oa.filter$citation.p.y,
  method = "spearman"
)

cor.test(
  DF.oa.filter$unique_country_codes,
  DF.oa.filter$citation.p.y,
  method = "spearman"
)
```


# Abstract
```{r}
# open OpenAlex data
DF.oa.filter <-
  readRDS(
    file.path(
      data_dir,
      "openalex",
      "btpsych24a_data_bib_processed.Rds"
    )
  )

# open Altmetrics data
DF.alt <-
  readRDS(
    file.path(
      data_dir,
      "altmetric",
      "btpsych24a_altmetric_processed.Rds"
    )
  )

# compute total number of citations
DF.oa.filter$cited_by_count %>% sum()

# compute total number of news mentions
DF.alt$news %>% sum()

# compute total number of social media mentions
DF.alt$gen %>% sum()

# compute total number of policy mentions
DF.alt$Policy.mentions %>% sum()
```

# Methodology
Total number of articles
```{r}
# open raw data
DF.oa.raw <-
  readRDS(
    file.path(
      data_dir,
      "openalex",
      "btpsych24a_data_bib_raw.Rds"
    )
  )

# total number of works compiled
DF.oa.raw %>%
  nrow()

# number of non-articles removed
DF.oa.raw %>%
  filter(
    type != "article"
  ) %>%
  nrow()

# number of articles without authorship info removed
DF.oa.raw %>%
  filter(
    type == "article",
    is.na(author)
  ) %>%
  nrow()

# number of duplicates
## pull out records with no NA's, as these will be labeled as 'duplicates'
DF.na <- DF.oa.raw %>%
  filter(
    type == "article",
    !is.na(author),
    is.na(doi)
  )

## remove duplicates
DF.oa.raw.nodup <-
  DF.oa.raw %>%
  filter(
    type == "article",
    !is.na(author)
  ) %>%
  distinct(doi,
    .keep_all = T
  )

## re-merge
DF.oa.raw.nodup <- rbind(DF.oa.raw.nodup, DF.na) %>%
  arrange(id)

## calculate number of duplicates that were removed
DF.oa.raw %>%
  filter(
    type == "article",
    !is.na(author)
  ) %>%
  nrow() - nrow(DF.oa.raw.nodup)

## calculate number of records removed based on frequently ineligible title phrases
DF.oa.raw.nodup %>%
  filter(
    !str_detect(
      title,
      regex("Transactions of|Transactions on|Conference|Annual meeting|Symposium|Proceedings of|Table of Contents",
        ignore_case = T
      )
    )
  ) %>%
  nrow() - nrow(DF.oa.raw.nodup)

# final openalex dataset
DF.oa.filter <-
  readRDS(
    file.path(
      data_dir,
      "openalex",
      "btpsych24a_data_bib_processed.Rds"
    )
  )

nrow(DF.oa.filter)
```

Indexing lag
```{r}
records.by.year <-
  DF.oa.raw %>%
  filter(type == "article") %>%
  group_by(publication_year) %>%
  summarise(n = n())

# decrease from 2020 to 2021
records.by.year[records.by.year$publication_year == "2020", ]$n -
  records.by.year[records.by.year$publication_year == "2021", ]$n

# decrease from 2020 to 2022
records.by.year[records.by.year$publication_year == "2020", ]$n -
  records.by.year[records.by.year$publication_year == "2022", ]$n

# decrease from 2020 to 2023
records.by.year[records.by.year$publication_year == "2020", ]$n -
  records.by.year[records.by.year$publication_year == "2023", ]$n
```

## Team size
Identify percentiles
```{r}
# rename dataframe so that you don't have to redo all the below code
DF <- DF.oa.filter

# calculate percent rank
per.DF <- DF %>%
  select(authors.n) %>%
  mutate(perc = percent_rank(authors.n) * 100)

# size.bin == 1 - 9
DF$bin.label <- NA

# create label for size.bin == 1 - 4
DF[DF$size.bin == "1 - 4", ]$bin.label <-
  paste0(
    per.DF[per.DF$authors.n == 1, ]$perc %>% max() %>% round(0),
    " - ",
    per.DF[per.DF$authors.n == 4, ]$perc %>% max() %>% round(1),
    " percentile"
  )

# create label for size.bin == 5 - 9
DF[DF$size.bin == "5 - 9", ]$bin.label <-
  paste0(
    per.DF[per.DF$authors.n == 5, ]$perc %>% max() %>% round(0),
    " - ",
    per.DF[per.DF$authors.n == 9, ]$perc %>% max() %>% round(1),
    " percentile"
  )

# size.bin == 10 - 19
DF[DF$size.bin == "10 - 19", ]$bin.label <-
  paste0(
    per.DF[per.DF$authors.n == 10, ]$perc %>% max() %>% round(1),
    " - ",
    per.DF[per.DF$authors.n == 19, ]$perc %>% max() %>% round(1),
    " percentile"
  )

# size.bin == 20 - 29
DF[DF$size.bin == "20 - 29", ]$bin.label <-
  paste0(
    per.DF[per.DF$authors.n == 20, ]$perc %>% max() %>% round(2),
    " - ",
    per.DF[per.DF$authors.n == 29, ]$perc %>% max() %>% round(2),
    " percentile"
  )

# size.bin == 30 - 39
DF[DF$size.bin == "30 - 39", ]$bin.label <-
  paste0(
    per.DF[per.DF$authors.n == 30, ]$perc %>% max() %>% round(3),
    " - ",
    per.DF[per.DF$authors.n == 39, ]$perc %>% max() %>% round(3),
    " percentile"
  )

# size.bin == 40 - 49
DF[DF$size.bin == "40 - 49", ]$bin.label <-
  paste0(
    per.DF[per.DF$authors.n == 40, ]$perc %>% max() %>% round(3),
    " - ",
    per.DF[per.DF$authors.n == 49, ]$perc %>% max() %>% round(3),
    " percentile"
  )

# size.bin == 50 - 59
DF[DF$size.bin == "50 - 59", ]$bin.label <-
  paste0(
    per.DF[per.DF$authors.n == 50, ]$perc %>% max() %>% round(3),
    " - ",
    per.DF[per.DF$authors.n == 59, ]$perc %>% max() %>% round(3),
    " percentile"
  )

# size.bin == 60 - 69
DF[DF$size.bin == "60 - 69", ]$bin.label <-
  paste0(
    per.DF[per.DF$authors.n == 60, ]$perc %>% max() %>% round(3),
    " - ",
    per.DF[per.DF$authors.n == 69, ]$perc %>% max() %>% round(3),
    " percentile"
  )

# size.bin == 70 - 79
DF[DF$size.bin == "70 - 79", ]$bin.label <-
  paste0(
    per.DF[per.DF$authors.n == 70, ]$perc %>% max() %>% round(3),
    " - ",
    per.DF[per.DF$authors.n == 79, ]$perc %>% max() %>% round(3),
    " percentile"
  )

# size.bin == 80 - 89
DF[DF$size.bin == "80 - 89", ]$bin.label <-
  paste0(
    per.DF[per.DF$authors.n == 80, ]$perc %>% max() %>% round(3),
    " - ",
    per.DF[per.DF$authors.n == 90, ]$perc %>% max() %>% round(3),
    " percentile"
  )

# size.bin == 90 - 99
DF[DF$size.bin == "90 - 99", ]$bin.label <-
  paste0(
    per.DF[per.DF$authors.n == 90, ]$perc %>% max() %>% round(3),
    " - ",
    per.DF[per.DF$authors.n == 99, ]$perc %>% max() %>% round(3),
    " percentile"
  )

# size.bin == 100 +
DF[DF$size.bin == "100 +", ]$bin.label <-
  paste0(
    " > ",
    per.DF[per.DF$authors.n == 100, ]$perc %>% max() %>% round(3),
    " percentile"
  )

# combine percentile and bin size information
DF <- DF %>%
  mutate(
    bin.label =
      paste0(
        size.bin, " co-authors", "\n",
        bin.label
      )
  )

# extract bin labels
bin.labels <- DF$bin.label %>%
  unique()

# illustrate concerns with stochastic variation
DF %>%
  filter(
    size.bin == "50 - 59",
    (publication_year == "2020" | publication_year == "2019")
  ) %>%
  group_by(publication_year) %>%
  summarise(n = n())

# calculate number and percentage of papers with 10 + authors
DF.oa.filter %>%
  filter(authors.n > 9) %>%
  nrow()

DF.oa.filter %>%
  filter(authors.n > 9) %>%
  nrow() / nrow(DF.oa.filter)

# delete vestigial
rm(per.DF, DF)
```

## Primary topics of inquiry
```{r}
# total number
DF.topics <-
  read.csv(
    file.path(
      data_dir,
      "topic",
      "btpsych24a_topic.csv"
    )
  )

nrow(DF.topics)

# reminder: not all size.bins had over 300 articles
DF %>%
  group_by(size.bin) %>%
  summarise(n = n())

# total number with topic classifications
DF.topics <- DF.topics %>%
  filter(
    topic != "N/A",
    topic != " N/A",
    topic != "N//A"
  )

nrow(DF.topics)
```

## Measures of impact
Number of mentions
```{r}
# compute total number of citations
DF.oa.filter$cited_by_count %>% sum()

# compute total number of news mentions
DF.alt$news %>% sum()

# compute total number of social media mentions
DF.alt$gen %>% sum()

# compute total number of policy mentions
DF.alt$Policy.mentions %>% sum()
```

Number of records searched in AltMetric DOI's
```{r}
rbind(
  read.csv(
    file.path(
      data_dir,
      "altmetric",
      "btpsych24a_bib_doi1.csv"
    )
  ) %>%
    select(doi),
  read.csv(
    file.path(
      data_dir,
      "altmetric",
      "btpsych24a_bib_doi2.csv"
    )
  ) %>%
    select(doi)
) %>%
  nrow()
```

Number of Altmetric data points
```{r}
DF.alt <-
  rbind(
    read.csv(
      file.path(
        data_dir,
        "altmetric",
        "btpsych24a_altmetric1_raw.csv"
      )
    ),
    read.csv(
      file.path(
        data_dir,
        "altmetric",
        "btpsych24a_altmetric2_raw.csv"
      )
    )
  )

nrow(DF.alt)
```

## Division of labor (supplemental analysis)
```{r}
# number of records included in coding
DF.c <-
  # open data
  read.csv(
    file.path(
      data_dir,
      "contributorship",
      "btpsych24a_cont_coding.csv"
    ),
    na.string = "N/A"
  )

nrow(DF.c)

# number of records actually coded
DF.c <- DF.c %>%
  filter(
    # remove instances without contributorship statements
    !is.na(contributorship.statement),
    contributorship.statement != "",

    # remove instances where coders recommended removal
    remove.revise != 1,

    # remove instances where there wasn't anything to code
    !(is.na(conception) &
      is.na(execution) &
      is.na(analysis) &
      is.na(draft.writing) &
      is.na(review.writing) &
      is.na(funding)
    )
  ) %>%
  # select relevant variables
  select(id, contributorship.statement:funding)

nrow(DF.c)
```

## Geographical distribution
```{r}
# number of big team efforts
DF.oa.filter %>%
  filter(authors.n > 9) %>%
  nrow()

# number of geocoded efforts (i.e., ones with complete authorship info)
DF.geo <- readRDS(
  file.path(
    data_dir,
    "location",
    "btpsych24a_geocode.Rds"
  )
)

DF.geo %>% length()
```

# Results
## Big team science is an emerging trend
Proportion of records published by teams of 10+ co-authors
```{r}
DF %>%
  filter(
    size.bin != "1 - 4",
    size.bin != "5 - 9"
  ) %>%
  nrow() / nrow(DF)
```

Growth rates
```{r}
DF %>%
  # focus on growth between 2011 and 2020
  filter(
    publication_year > 2010,
    publication_year < 2021
  ) %>%
  # distinguish between first and second half of that decade
  mutate(
    pub.period =
      if_else(publication_year > 2015,
        "later",
        "earlier"
      )
  ) %>%
  # calculate change over time for each size.bin
  group_by(size.bin, pub.period) %>%
  summarise(n = n()) %>%
  pivot_wider(
    names_from = "pub.period",
    values_from = n,
    id_cols = "size.bin"
  ) %>%
  rowwise() %>%
  mutate(
    per.growth = (later - earlier) / earlier,
    per.growth = per.growth * 100,
    per.growth = round(per.growth, 3)
  ) %>%
  View()
```

Big team science by topic
```{r}
# recode topic
DF.topics <- DF.topics %>%
  mutate(
    topic.trunc =
      substr(topic, 1, 2),
    topic.trunc =
      recode(topic.trunc,
        "21" = "General Psychology",
        "22" = "Psychometrics & Statistics & Methodology",
        "23" = "Human Experimental Psychology",
        "24" = "Animal Experimental & Comparative Psychology",
        "25" = "Physiological Psychology & Neuroscience",
        "26" = "Psychology & the Humanities",
        "27" = "Communication Systems",
        "28" = "Developmental Psychology",
        "29" = "Social Processes & Social Issues",
        "30" = "Social Psychology",
        "31" = "Personality Psychology",
        "32" = "Psychological & Physical Disorders",
        "33" = "Health & Mental Health Treatment & Prevention",
        "34" = "Health & Mental Health Personnel Issues",
        "35" = "Educational & School Psychology",
        "36" = "Organizational Psychology & Human Resources",
        "37" = "Sport Psychology & Leisure",
        "38" = "Military Psychology",
        "39" = "Consumer Psychology",
        "40" = "Engineering & Environmental Psychology",
        "41" = "Cognitive Psychology & Intelligent Systems",
        "42" = "Forensic Psychology & Legal Issues"
      )
  )

DF.topics %>%
  # focus on big team efforts
  filter(size.bin != "1 - 4") %>%
  # group by topic category
  group_by(topic.trunc) %>%
  # summarise and organize
  summarise(n = n()) %>%
  mutate(
    prop = n / sum(n),
    prop = prop * 100
  ) %>%
  arrange(-prop) %>%
  View()
```

## Bigger teams produce work...
Upload impact data
```{r}
impact.by.size <-
  readRDS(here("figures", "impact.viz.DF.Rds"))

DF.alt <- readRDS(
  file.path(
    data_dir,
    "altmetric",
    "btpsych24a_altmetric_processed.Rds"
    )
  )
```

1-4 co-authors
```{r}
# proportion of available records
impact.by.size[impact.by.size$size.bin == "1 - 4", ]$n[1] /
  sum(impact.by.size$n,
    na.rm = T
  ) * 100

# stats
impact.by.size[impact.by.size$size.bin == "1 - 4", ]
```

10-19 co-authors
```{r}
# proportion of available records
impact.by.size[impact.by.size$size.bin == "10 - 19", ]$n[1] /
  sum(impact.by.size$n,
    na.rm = T
  ) * 100

# stats
impact.by.size[impact.by.size$size.bin == "10 - 19", ]

#####################
# citation comparison
#####################
3.36 / 0.927

mes(
  n.2 = impact.by.size[9, ]$n,
  m.2 = impact.by.size[9, ]$m,
  sd.2 = impact.by.size[9, ]$sd,
  n.1 = impact.by.size[1, ]$n,
  m.1 = impact.by.size[1, ]$m,
  sd.1 = impact.by.size[1, ]$sd
)

ttestBF(x = DF[DF$size.bin == '1 - 4', ]$citation.p.y,
        y = DF[DF$size.bin == '10 - 19', ]$citation.p.y)

#####################
# news comparison
#####################
0.62 / 0.191

# cohen's d
mes(
  n.2 = impact.by.size[9, ]$n,
  m.2 = impact.by.size[11, ]$m,
  sd.2 = impact.by.size[11, ]$sd,
  n.1 = impact.by.size[1, ]$n,
  m.1 = impact.by.size[3, ]$m,
  sd.1 = impact.by.size[3, ]$sd
)

ttestBF(x = DF.alt[DF.alt$size.bin == '1 - 4' & 
                     !is.na(DF.alt$news.p.y), ]$news.p.y,
        y = DF.alt[DF.alt$size.bin == '10 - 19' & 
                     !is.na(DF.alt$news.p.y), ]$news.p.y)

#####################
# social media comparison
#####################
3.62 / 1.43

mes(
  n.2 = impact.by.size[9, ]$n,
  m.2 = impact.by.size[12, ]$m,
  sd.2 = impact.by.size[12, ]$sd,
  n.1 = impact.by.size[1, ]$n,
  m.1 = impact.by.size[4, ]$m,
  sd.1 = impact.by.size[4, ]$sd
)

ttestBF(x = DF.alt[DF.alt$size.bin == '1 - 4' & 
                     !is.na(DF.alt$gen.p.y), ]$gen.p.y,
        y = DF.alt[DF.alt$size.bin == '10 - 19' & 
                     !is.na(DF.alt$gen.p.y), ]$gen.p.y)

#####################
# policy comparison
#####################
0.0179 / 0.00758

mes(
  n.2 = impact.by.size[9, ]$n,
  m.2 = impact.by.size[10, ]$m,
  sd.2 = impact.by.size[10, ]$sd,
  n.1 = impact.by.size[1, ]$n,
  m.1 = impact.by.size[2, ]$m,
  sd.1 = impact.by.size[2, ]$sd
)

ttestBF(x = DF.alt[DF.alt$size.bin == '1 - 4' & 
                     !is.na(DF.alt$policy.p.y), ]$policy.p.y,
        y = DF.alt[DF.alt$size.bin == '10 - 19' & 
                     !is.na(DF.alt$policy.p.y), ]$policy.p.y)
```

30-39 co-authors
```{r}
# proportion of available records
impact.by.size[impact.by.size$size.bin == "30 - 39", ]$n[1] /
  sum(impact.by.size$n,
    na.rm = T
  ) * 100

# stats
impact.by.size[impact.by.size$size.bin == "30 - 39", ] %>% View()

#####################
# citation comparison
#####################
8.90 / 0.927
mes(
  n.2 = impact.by.size[17, ]$n,
  m.2 = impact.by.size[17, ]$m,
  sd.2 = impact.by.size[17, ]$sd,
  n.1 = impact.by.size[1, ]$n,
  m.1 = impact.by.size[1, ]$m,
  sd.1 = impact.by.size[1, ]$sd
)

ttestBF(x = DF[DF$size.bin == '1 - 4', ]$citation.p.y,
        y = DF[DF$size.bin == '30 - 39', ]$citation.p.y)

#####################
# news comparison
#####################
3.1 / 0.191
mes(
  n.2 = impact.by.size[17, ]$n,
  m.2 = impact.by.size[19, ]$m,
  sd.2 = impact.by.size[19, ]$sd,
  n.1 = impact.by.size[1, ]$n,
  m.1 = impact.by.size[3, ]$m,
  sd.1 = impact.by.size[3, ]$sd
)

ttestBF(x = DF.alt[DF.alt$size.bin == '1 - 4' & 
                     !is.na(DF.alt$news.p.y), ]$news.p.y,
        y = DF.alt[DF.alt$size.bin == '30 - 39' & 
                     !is.na(DF.alt$news.p.y), ]$news.p.y)

#####################
# social media comparison
19.9 / 1.43
#####################
mes(
  n.2 = impact.by.size[17, ]$n,
  m.2 = impact.by.size[20, ]$m,
  sd.2 = impact.by.size[20, ]$sd,
  n.1 = impact.by.size[1, ]$n,
  m.1 = impact.by.size[4, ]$m,
  sd.1 = impact.by.size[4, ]$sd
)

ttestBF(x = DF.alt[DF.alt$size.bin == '1 - 4' & 
                     !is.na(DF.alt$gen.p.y), ]$gen.p.y,
        y = DF.alt[DF.alt$size.bin == '30 - 39' & 
                     !is.na(DF.alt$gen.p.y), ]$gen.p.y)

#####################
# policy comparison
#####################
0.0636 / 0.00758

mes(
  n.2 = impact.by.size[17, ]$n,
  m.2 = impact.by.size[18, ]$m,
  sd.2 = impact.by.size[18, ]$sd,
  n.1 = impact.by.size[1, ]$n,
  m.1 = impact.by.size[2, ]$m,
  sd.1 = impact.by.size[2, ]$sd
)

ttestBF(x = DF.alt[DF.alt$size.bin == '1 - 4' & 
                     !is.na(DF.alt$policy.p.y), ]$policy.p.y,
        y = DF.alt[DF.alt$size.bin == '30 - 39' & 
                     !is.na(DF.alt$policy.p.y), ]$policy.p.y)
```

Correlations
```{r}
DF.impact <-
  readRDS(here(
    "figures",
    "impact.viz.DF.raw.Rds"
  ))

cor.test(DF.impact$authors.n,
  DF.impact$citation.p.y,
  method = "spearman"
)

cor.test(DF.impact$authors.n,
  DF.impact$news.p.y,
  method = "spearman"
)

cor.test(DF.impact$authors.n,
  DF.impact$gen.p.y,
  method = "spearman"
)

cor.test(DF.impact$authors.n,
  DF.impact$policy.p.y,
  method = "spearman"
)

## Bayesian correlation (Bayes Factor) analogs to Spearman tests
d <- DF.impact %>%
  mutate(
    citation.p.y = rank(citation.p.y),
    news.p.y = rank(news.p.y),
    gen.p.y = rank(gen.p.y),
    policy.p.y = rank(policy.p.y),
    authors.n = rank(authors.n),
  ) %>%
  data.frame()

lmBF(citation.p.y ~ authors.n,
  data = d
)

lmBF(news.p.y ~ authors.n,
  data = d
)

lmBF(gen.p.y ~ authors.n,
  data = d
)

lmBF(policy.p.y ~ authors.n,
  data = d
)
```

## Bayesian linear vs. quadratic comparisons (minimal)
```{r bayes-linear-quad-minimal, message=FALSE, warning=FALSE}
# Minimal Bayesian analysis: compare linear vs quadratic relationships
# across per-year impact measures using brms. We use standard weakly
# informative priors on standardized predictors and model the same
# per-year outcomes used in the frequentist analyses.
library(brms)
library(bridgesampling)
library(loo)
library(bayestestR)

# Ensure DF.100 exists (created in the frequentist section)
if (!exists("DF.100")) {
  DF.100 <- DF.impact %>% filter(authors.n < 100)
}

DF.100 <- DF.100 %>%
  mutate(
    a_s = as.numeric(scale(authors.n)),
    a_s2 = a_s^2
  )

# Standard weakly-informative priors for standardized predictors
priors_standard <- c(
  prior(normal(0, 1), class = "b"), # slope
  prior(normal(0, 5), class = "Intercept"), # intercept
  prior(student_t(3, 0, 2.5), class = "sigma")
)

outcomes <- c("citation.p.y", "news.p.y", "gen.p.y", "policy.p.y")

bayes_results <- lapply(outcomes, function(out) {
  cat("\n---\nOutcome: ", out, "\n")
  
  f_lin <- brms::bf(as.formula(paste0(out, " ~ a_s")))
  f_quad <- brms::bf(as.formula(paste0(out, " ~ a_s + a_s2")))

  m_lin <- brm(f_lin,
    data = DF.100,
    family = gaussian(),
    prior = priors_standard
  )

  m_quad <- brm(f_quad,
    data = DF.100,
    family = gaussian(),
    prior = priors_standard
  )

  # bridge sampling for marginal likelihoods -> Bayes factor (quad vs lin)
  bs_lin <- bridge_sampler(m_lin, silent = TRUE)
  bs_quad <- bridge_sampler(m_quad, silent = TRUE)
  bf_qvsl <- bridgesampling::bf(bs_quad, bs_lin)

  list(
    m_lin = m_lin,
    m_quad = m_quad,
    bf = bf_qvsl
  )
})

# Set names for the list elements
names(bayes_results) <- outcomes

# export bayes_results
saveRDS(bayes_results, here("figures", "bayes_linear_quad_results.Rds"))

bayes_results$policy.p.y$bf

# Brief posterior checks for the last fitted model
pp_check(m_quad, ndraws = 30)
print(bayes_R2(m_quad))
```

### Comparisons of linear vs. quadratic relationships
Prepare for separate analyses for collaborations (a) up to 100 and (b) beyond 100
```{r}
# focus dataframe on less than 100
DF.100 <- DF.impact %>%
  filter(authors.n < 100)
```

#### Impact
```{r}
m.l <-
  lm(citation.p.y ~ authors.n,
    data = DF.100
  )

m.q <-
  lm(citation.p.y ~ authors.n + I(authors.n^2),
    data = DF.100
  )

anova(m.l, m.q)
AIC(m.l, m.q)$AIC %>%
  diff()

rm(m.l, m.q)

ggplot(
  DF.impact %>% filter(authors.n < 100),
  aes(x = authors.n, y = policy.p.y)
) +
  # model linear relationship
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    color = "grey50",
    linetype = "dashed",
    linewidth = .75
  ) +

  # model quadratic relationsnip
  geom_smooth(
    method = "lm",
    formula = y ~ x + I(x^2),
    color = "grey50",
    linewidth = .75
  )
```

#### News
```{r}
m.l <-
  lm(news.p.y ~ authors.n,
    data = DF.100
  )

m.q <-
  lm(news.p.y ~ authors.n + I(authors.n^2),
    data = DF.100
  )

anova(m.l, m.q)
AIC(m.l, m.q)$AIC %>%
  diff()

rm(m.l, m.q)

ggplot(
  DF.impact %>% filter(authors.n < 100),
  aes(x = authors.n, y = news.p.y)
) +
  # model linear relationship
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    color = "grey50",
    linetype = "dashed",
    linewidth = .75
  ) +

  # model quadratic relationsnip
  geom_smooth(
    method = "lm",
    formula = y ~ x + I(x^2),
    color = "grey50",
    linewidth = .75
  )
```

#### Social media
```{r}
m.l <-
  lm(gen.p.y ~ authors.n,
    data = DF.100
  )

m.q <-
  lm(gen.p.y ~ authors.n + I(authors.n^2),
    data = DF.100
  )

anova(m.l, m.q)
AIC(m.l, m.q)$AIC %>%
  diff()

rm(m.l, m.q)

ggplot(
  DF.impact %>% filter(authors.n < 100),
  aes(x = authors.n, y = gen.p.y)
) +
  # model linear relationship
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    color = "grey50",
    linetype = "dashed",
    linewidth = .75
  ) +

  # model quadratic relationsnip
  geom_smooth(
    method = "lm",
    formula = y ~ x + I(x^2),
    color = "grey50",
    linewidth = .75
  )
```

#### Policy
```{r}
m.l <-
  lm(policy.p.y ~ authors.n,
    data = DF.100
  )

m.q <-
  lm(policy.p.y ~ authors.n + I(authors.n^2),
    data = DF.100
  )

anova(m.l, m.q)

AIC(m.l, m.q)$AIC %>%
  diff()

AIC
rm(m.l, m.q)
```

## Sensitivity analyses
### Linear and non-linear regression with massive team size outliers included
```{r}
# number of papers with over 100 articles
DF.impact %>%
  filter(authors.n > 99) %>%
  nrow()

DF.impact$authors.n %>% max()

DF.impact.full <- DF.impact %>%
  rename(
    "scholarly articles" = "citation.p.y",
    "news" = "news.p.y",
    "social media" = "gen.p.y",
    "policy documents" = "policy.p.y"
  ) %>%
  pivot_longer("scholarly articles":"social media")

ggplot(
  data = DF.impact.full,
  aes(
    x = authors.n,
    y = value
  )
) +
  facet_wrap(~name,
    scales = "free"
  ) +
  # model linear relationship
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    color = "grey50",
    linetype = "dashed",
    linewidth = .75
  ) +

  # model quadratic relationsnip
  geom_smooth(
    method = "lm",
    formula = y ~ x + I(x^2),
    color = "grey50",
    linewidth = .75
  )
```

### Alternative statistical operationalization of impact (OpenAlex)
```{r}
oa.sens <- DF.oa.filter %>%
  # examine whether record received any attention
  mutate(
    scholar.att =
      if_else(cited_by_count > 0,
        1,
        0
      )
  ) %>%
  # group by size bin
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  # calculate relative summary statistics
  summarise(
    mn.yearly_scholar = mean(citation.p.y) %>%
      round(2),
    md.yearly_scholar = median(citation.p.y) %>%
      round(2),
    mn_scholar = mean(cited_by_count) %>%
      round(2),
    md_scholar = median(cited_by_count) %>%
      round(2),
    att_scholar = mean(scholar.att) %>%
      round(2)
  )
```

### Alternative statistical operationalization of impact Altmetric
```{r}
# open Altmetric data
DF.alt <- readRDS(
  file.path(
    data_dir,
    "altmetric",
    "btpsych24a_altmetric_processed.Rds"
  )
)
```

```{r}
# sensitivity analyses
alt.sens <- DF.alt %>%
  # examine whether record received any attention
  mutate(
    policy.att =
      if_else(Policy.mentions > 0,
        1,
        0
      ),
    news.att =
      if_else(news > 0,
        1,
        0
      ),
    gen.att =
      if_else(gen > 0,
        1,
        0
      )
  ) %>%
  # group by size bin
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  # calculate relative summary statistics
  summarise(
    mn.yearly_news = mean(news.p.y) %>%
      round(2),
    md.yearly_news = median(news.p.y) %>%
      round(2),
    mn_news = mean(news) %>%
      round(2),
    md_news = median(news) %>%
      round(2),
    att_news = mean(news.att) %>%
      round(2),
    mn.yearly_social = mean(gen.p.y) %>%
      round(2),
    md.yearly_social = median(gen.p.y) %>%
      round(2),
    mn_social = mean(gen) %>%
      round(2),
    md_social = median(gen) %>%
      round(2),
    att_social = mean(gen.att) %>%
      round(2),
    mn.yearly_policy = mean(policy.p.y) %>%
      round(2),
    md.yearly_policy = median(policy.p.y) %>%
      round(2),
    mn_policy = mean(Policy.mentions) %>%
      round(2),
    md_policy = median(Policy.mentions) %>%
      round(2),
    att_policy = mean(policy.att) %>%
      round(2)
  )

sens <-
  full_join(oa.sens,
    alt.sens,
    by = "size.bin"
  )
```


```{r}
## Overall, median
DF.alt %>%
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  summarise(
    med.policy = median(Policy.mentions),
    med.news = median(news),
    med.gen = median(gen)
  )

## Yearly, median
DF.alt %>%
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  summarise(
    med.policy = median(policy.p.y),
    med.news = median(news.p.y),
    med.gen = median(gen.p.y)
  )

## Likelihood of being mentioned
DF.alt %>%
  mutate(
    Policy.mentions =
      if_else(Policy.mentions > 0,
        1,
        0
      ),
    news =
      if_else(news > 0,
        1,
        0
      ),
    gen =
      if_else(gen > 0,
        1,
        0
      )
  ) %>%
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  summarise(
    prop.policy = mean(Policy.mentions),
    prop.news = mean(news),
    prop.gen = mean(gen)
  )
```

## Big teams are mobilized (supplemental analysis)
```{r}
DF.c <- readRDS(
  file.path(
    data_dir,
    "contributorship",
    "btpsych24a_cont_processed.Rds"
  )
)

# proportion
DF.c %>%
  mutate(prop.n = prop.n * 100) %>%
  group_by(category) %>%
  summarise(
    m = mean(prop.n,
      na.rm = T
    ),
    sd = sd(prop.n,
      na.rm = T
    )
  )

# number that focused on each category
DF.c %>%
  filter(dominant == "yes") %>%
  group_by(category.main) %>%
  summarise(prop = (n() / nrow(DF.c)) * 100) %>%
  mutate(other = 100 - sum(prop))

# variability within categories
DF.c %>%
  filter(
    category.main == "conceptualization focused (14%)",
    dominant == "yes",
    category == "conception"
  ) %>%
  summarise(
    min = min(prop.n, na.rm = T) * 100,
    max = max(prop.n, na.rm = T) * 100
  )
```


## There are economic...
First author positions
```{r}
# open data
f.author <-
  read.csv(file = file.path(
    data_dir,
    "worlddata",
    "btpsych24a_f.author_gdp.csv"
  ))

# look at top contributors
f.author %>%
  mutate(
    prop.n = n / sum(n),
    prop.n = prop.n * 100
  ) %>%
  arrange(desc(prop.n)) %>%
  .[1:5, ] %>%
  mutate(total = sum(prop.n))

# correlations with gdp
cor.test(
  f.author$n,
  f.author$m.gdp.pc
)

cor.test(
  f.author$n,
  f.author$m.gdp
)
`
``
  
Non-first author positions
```{r}
# open data
s.author <-
  read.csv(
    file = file.path(
      data_dir,
      "worlddata",
      "btpsych24a_s.author_gdp.csv"
    )
  )

# look at top contributors
s.author %>%
  mutate(
    prop.n = n / sum(n),
    prop.n = prop.n * 100
  ) %>%
  arrange(desc(prop.n)) %>%
  .[1:5, ] %>%
  mutate(total = sum(prop.n))

# correlations with gdp
cor.test(
  s.author$n,
  s.author$m.gdp.pc
)

cor.test(
  s.author$n,
  s.author$m.gdp
)
```
