---
title: "Create map figure"
author: "Anonymized for peer review (NC)"
reviewer: "Anonymized for peer review (JT)"
output: word_document
editor_options: 
  chunk_output_type: console
---

This code creates a figure that (a) illustrates the geographical distribution of big team leadership and participation rates in psychology, and (b) examines links between those rates and indicators of economic wealth (GDP)

# Setup
```{r setup, include= F}
# specify data directory
data_dir <- Sys.getenv("data_dir")

# source setup file
source('setup.R')
```

# open data files
```{r}
DF.d <- 
  read_csv(file.path(data_dir,
                    "worlddata",
                    "demo",
                    "btpsych24a_demo.csv"),
           show_col_types = F
          )

DF.e <- 
  read_csv(file.path(data_dir,
                    "worlddata",
                    "edu",
                    "btpsych24a_edu.csv"),
           show_col_types = F,
           skip = 4
          )

DF.r <- 
  read_csv(file.path(data_dir,
                    "worlddata",
                    "gdp.pc",
                    "btpsych24a_worlddata_gdp.pc.csv"),
           show_col_types = F,
           skip = 4
          )

DF.i <- 
  read_csv(file.path(data_dir,
                    "worlddata",
                    "gdp.sector",
                    "btpsych24a_gdp.sector.csv"),
           show_col_types = F,
           skip = 4
          )
```



# Prepare OpenAlex data
## Open and isolate OpenAlex authorship data
```{r open.alex, include = F}
DF.author <- 
  # open data
  readRDS(file.path(data_dir,
                    "openalex",
                    "btpsych24a_data_bib_processed.Rds")) %>% 
  
  # isolate big team science records
  filter(size.bin != "1 - 4",
         size.bin != "5 - 9")

# focus on author lists
DF.author <- DF.author$author
```

## Geocode records
```{r}
DF.geo <- 
  # isolate affiliation information
  lapply(DF.author,
         function(sublist)
           c(sublist['institution_display_name'],
             sublist['institution_country_code'])
         ) %>% 
  
  # select only instances with full affiliation information
  ## changed to *any insatance* on Jul 28 2025 following reviewer recommendation
  ### this drops number of records from 49919 to 41107
  #Filter(function(sublist) all(!is.na(unlist(sublist))), .) %>% 
  Filter(function(sublist) any(!is.na(unlist(sublist))), .) %>% 
  
  # geocode using institution name and country
  lapply(.,
         function(sublist)
           as_tibble(sublist) %>% 
           mutate(address = paste0(institution_display_name,
                                   ", ",
                                   institution_country_code)
                  ) %>% 
           tidygeocoder::geocode(.,
                                 address = address,
                                 method = 'arcgis', 
                                 lat = latitude, 
                                 long = longitude
                                 )
         )

# save geocoded list
saveRDS(DF.geo,
        file.path(data_dir,
                  'location',
                  'btpsych24a_geocode.Rds')
        )
```

## Identify leaders and supporters
```{r}
# reload data (if needed)
DF.geo <- 
  readRDS(here('data',
               'location',
               'btpsych24a_geocode.Rds')
          )

DF.geo <-
  
  # separate leader and supporter location information
  lapply(DF.geo,
         function(sublist)
           sublist %>% 
           mutate(
             leader_coun = .[['institution_country_code']][1],
             leader_lat = .[["latitude"]][1],
             leader_lon = .[["longitude"]][1],
             supporter_coun = .[['institution_country_code']],
             supporter_lat = .[["latitude"]],
             supporter_lon = .[["longitude"]]
           )
         )
```

## Identify regions
```{r}
DF.geo <- 
  lapply(DF.geo,
         function(sublist)
           sublist %>% 
           mutate(
             # leader region
             leader_region = countrycode(.[['leader_coun']], 
                                         origin = 'iso2c', 
                                         destination = 'un.regionsub.name'),
             # supporter region
             supporter_region = countrycode(.[['supporter_coun']], 
                                            origin = 'iso2c', 
                                            destination = 'un.regionsub.name')
           )
         )
```


# Prepare GDP figures

## Load GDP and GDP per capita data
```{r}
# raw GDP
DF.gdp.raw <- 
  read.csv(here('data', 
                'worlddata', 
                'btpsych24a_worlddata_gdp.csv'),
           skip = 4) %>% # skip 4 empty rows
  select(Country.Code, X2023) %>% 
  rename("iso_a3" = Country.Code,
         "gdp" = X2023)

# GDP per capita
DF.gdp.pc <-
    read.csv(here('data', 
                  'worlddata', 
                  'btpsych24a_worlddata_gdp.pc.csv'),
           skip = 4) %>%
  select(Country.Code, X2023) %>% 
  rename("iso_a3" = Country.Code,
         "gdp.pc" = X2023)

# merge
DF.gdp <- left_join(DF.gdp.raw,
                    DF.gdp.pc,
                    by = "iso_a3")

# open ISO identifiers
world <- ne_countries(scale = "medium", 
                      returnclass = "sf")

# fix missing ISO values in world data
world[world$name == "France", ]$iso_a2 = "FR"
world[world$name == "France", ]$iso_a3 = "FRA"
world[world$name == "Norway", ]$iso_a2 = "NO"
world[world$name == "Norway", ]$iso_a3 = "NOR"
world[world$name == "Norway", ]$iso_a3 = "NOR"
world[world$name == "Kosovo", ]$iso_a2 = "XK"
world[world$name == "Kosovo", ]$iso_a3 = "XKX"

# select ISO identifiers
world <- world %>% 
  select(iso_a2, iso_a3)

# connect ISO identifiers to GDP data
DF.gdp <- left_join(DF.gdp, 
                    world,
                    by = "iso_a3")

# delete vestigial
rm(DF.gdp.pc, DF.gdp.raw)
```

## Examine links with first author positions
### Prepare and export data
```{r}
# prepare data
f.author <-
  # isolate first authors
  lapply(DF.geo,
         function(sublist)
           slice(sublist,
                 1)
         ) %>% 
  
  # bind rows
  bind_rows() %>% 
  
  # select country information
  select(leader_coun, leader_region) %>% 
  rename(iso_a2 = leader_coun) %>% 
  
  # join with GDP data
  left_join(.,
            DF.gdp,
            by = "iso_a2") %>% 
  
  # calculate mean GDP and number of records
  group_by(iso_a2) %>% 
  summarise(region = unique(leader_region),
            m.gdp = mean(gdp,
                         na.rm = T),
            m.gdp = m.gdp / 100000000, ## transformed for easier visualization
            m.gdp.pc = mean(gdp.pc,
                            na.rm = T),
            n = n()) %>%
  ungroup()

# export data
write.csv(x = f.author,
          file = here('data',
                      'worlddata',
                      'btpsych24a_f.author_gdp.csv'),
          row.names = F)
```

### Create figures
```{r}
# load data (if necessary)
f.author <- 
  read.csv(file = here('data',
                       'worlddata',
                       'btpsych24a_f.author_gdp.csv')
           )

# create GDP per capita plot
gdp.pc.lead <- 
  ggplot(data = f.author,
         aes(x = m.gdp.pc,
             y = n)) +
  
  # plot points and smoothed line
  geom_smooth(method = 'lm',
              fill = 'grey90',
              color = "grey70",
              alpha = .5) +
  geom_point(aes(color = region),
             alpha = 0.4,
             size = 2,
             show.legend = F) +
  
  # aesthetics
  xlab("GDP per capita") +
  ylab("First authored papers")

# create GDP plot
gdp.lead <- 
  ggplot(data = f.author,
         aes(x = m.gdp,
             y = n)) +
  
  # plot points and smoothed line
  geom_smooth(method = 'lm',
              fill = 'grey90',
              color = "grey70",
              alpha = .5) +
  geom_point(aes(color = region),
             alpha = 0.4,
             size = 2,
             show.legend = F) +
  # aesthetics
  theme(axis.title.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
        ) +
  xlab("GDP")

# combine plots
gdp.lead.fig <- ggdraw() +
  draw_plot(gdp.pc.lead) +
  draw_plot(gdp.lead, 
            x = .17,
            y = .7,
            width = .25,
            height = .25)

rm(gdp.pc.lead, gdp.lead, f.author)
```

## Examine links with non-first author positions
### Prepare and export data
```{r}
# prepare data
s.author <-
  # isolate first authors
  lapply(DF.geo,
         function(sublist)
           slice(sublist,
                 -1)
         ) %>% 
  
  # bind rows
  bind_rows() %>% 
  
  # select country information
  select(supporter_coun, supporter_region) %>% 
  rename(iso_a2 = supporter_coun) %>% 
  
  # join with GDP data
  left_join(.,
            DF.gdp,
            by = "iso_a2") %>% 
  
  # calculate mean GDP and number of records
  group_by(iso_a2) %>% 
  summarise(region = unique(supporter_region),
            m.gdp = mean(gdp,
                         na.rm = T),
            m.gdp = m.gdp / 100000000,
            m.gdp.pc = mean(gdp.pc,
                            na.rm = T),
            n = n()) %>%
  ungroup()

# export data
write.csv(x = s.author,
          file = here('data',
                      'worlddata',
                      'btpsych24a_s.author_gdp.csv'),
          row.names = F)
```

### Create figures
```{r}
# load data (if necessary)
s.author <- 
  read.csv(file = here('data',
                       'worlddata',
                       'btpsych24a_s.author_gdp.csv')
           )

# create GDP per capita plot
gdp.pc.support <- 
  ggplot(data = s.author,
         aes(x = m.gdp.pc,
             y = n)) +
  
  # plot points and smoothed line
  geom_smooth(method = 'lm',
              fill = 'grey90',
              color = "grey70",
              alpha = .5) +
  geom_point(aes(color = region),
             alpha = 0.4,
             size = 2,
             show.legend = F) +
  
  # aesthetics
  xlab("GDP per capita") +
  ylab("Non-first authored papers")

# create GDP plot
gdp.support <- 
  ggplot(data = s.author,
         aes(x = m.gdp,
             y = n)) +
  
  # plot points and smoothed line
  geom_smooth(method = 'lm',
              fill = 'grey90',
              color = "grey70",
              alpha = .5) +
  geom_point(aes(color = region),
             alpha = 0.4,
             size = 2,
             show.legend = F) +
  
  # aesthetics
  theme(axis.title.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
        ) +
  xlab("GDP")

# combine plots
gdp.support.fig <- ggdraw() +
  draw_plot(gdp.pc.support) +
  draw_plot(gdp.support, 
            x = .17,
            y = .7,
            width = .25,
            height = .25)

rm(gdp.pc.support, gdp.support, s.author)
```

## Combine 
```{r}
gdp.fig <- plot_grid(gdp.lead.fig, gdp.support.fig)

rm(gdp.lead.fig, gdp.support.fig)
```

# Prepare globe figures
```{r}
DF.map <-
  
  # remove first author from geocoded data (we will only plot points for supporters)
  lapply(DF.geo,
         function(sublist)
           slice(sublist,
                 -1)
         ) %>% 
  
  # bind_rows
  bind_rows() %>% 
  
  #select relevant variables
  select(leader_coun : supporter_lon) %>% 
  
  # filter out instances where collaborators were in the same location
  filter(paste0(leader_lat, ", ", leader_lon) !=
           paste0(supporter_lat, ", ", supporter_lon)) %>% 
  
  # filter out incomplete observations
  filter(!is.na(supporter_lat),
         !is.na(leader_lat)) %>% 
  
  # identify subregion name
  mutate(leader.region = 
           countrycode(.$leader_coun, 
                       origin = 'iso2c', 
                       destination = 'un.regionsub.name'),
         supporter.region = 
           countrycode(.$supporter_coun, 
                       origin = 'iso2c', 
                       destination = 'un.regionsub.name')
         )

# manually fix coding for TW and XK
DF.map[DF.map$leader_coun == "TW", ]$leader.region <- "Eastern Asia"
DF.map[DF.map$supporter_coun == "TW", ]$supporter.region <- "Eastern Asia"
DF.map[DF.map$leader_coun == "XK", ]$leader.region <- "Eastern Europe"
DF.map[DF.map$supporter_coun == "XK", ]$supporter.region <- "Eastern Europe"

# transform coordinates to crs 4326 map projection
## Note: this needs to be done separately for supporter and leader data
supporter.coord <- DF.map %>% 
  select(supporter.region, supporter_lat, supporter_lon) %>% 
  st_as_sf(x = .,
           coords = c('supporter_lon', 
                      'supporter_lat'),
           crs = 4326,
           remove = F) %>% 
  st_transform(., crs = "ESRI:54030") %>% 
  mutate(supporter_lat_new = st_coordinates(.)[, 2],
         supporter_lon_new = st_coordinates(.)[,1])

leader.coord <- DF.map %>% 
    select(leader.region, leader_lat, leader_lon) %>% 
  st_as_sf(x = .,
           coords = c('leader_lon', 
                      'leader_lat'),
           crs = 4326,
           remove = F) %>% 
  st_transform(., crs = "ESRI:54030") %>% 
  mutate(leader_lat_new = st_coordinates(.)[, 2],
         leader_lon_new = st_coordinates(.)[,1])

DF.map <- cbind(supporter.coord, leader.coord)

rm(supporter.coord, leader.coord)
```

## Globe figure
```{r}
globe.fig <- ggplot() +
  # plot map layer
  layer_spatial(data = world,,
                color = NA)+
  
  # change coordinate projections
  coord_sf(crs = 'ESRI:54030') +
  
  # add dots for supporters
  geom_point(data = DF.map, 
             aes(x = supporter_lon_new, 
                 y = supporter_lat_new,
                 colour = supporter.region),
             size = .4,
             alpha = .01) +
  
  # connect leaders and supporters
  geom_curve(data = DF.map, 
             aes(x = leader_lon_new, 
                 y = leader_lat_new, 
                 xend = supporter_lon_new, 
                 yend = supporter_lat_new,
                 colour = leader.region),
             curvature = -0.5,
             size = 0.1, 
             alpha = .01,
             show.legend = F) +
  
  # improve aesthetics
  cowplot::theme_minimal_grid() +
  
  theme(legend.position = 'bottom',
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.justification = "center",
        panel.grid.major = element_line(size = .3)) +
  
  guides(color = 
           guide_legend(override.aes = list(size = 3,
                                            alpha = 1),
                        ncol = 4)) +
  
  labs(x = NULL, y = NULL)
```

# Combine globe and gdp figures
```{r}
# combine globe and gdp figures
final.fig <- 
  plot_grid(globe.fig,
            gdp.fig,
            nrow = 2,
            rel_heights = c(6, 4),
            labels = c("A", "B"))
```
