---
title: "Create map figure"
output: word_document
editor_options: 
  chunk_output_type: console
---

# Setup
```{r setup, include= F}
# specify data directory
data_dir <- Sys.getenv("data_dir")

# source setup file
source("setup.R")
```

# Open bibliometric data and identify locations
```{r}
# open data
DF.o <-
  # open data
  readRDS(file.path(
    data_dir,
    "openalex",
    "btpsych24a_data_bib_processed.Rds"
  ))

# identify country of first author via indexing
DF.o$author1.country <-
  lapply(
    c(1:nrow(DF.o)),
    function(x) {
      DF.o[x, ][["author"]][[1]][["institution_country_code"]][[1]] %>%
        return()
    }
  )

# Prepare for merging
DF.o <- DF.o %>%
  mutate(
    iso3 =
      author1.country %>%
        # convert to character and identify NAs
        as.character() %>%
        if_else(. == "NA",
          NA,
          .
        ) %>%
        # convert country code iso3 format (to facilitate later merge)
        countrycode(.,
          origin = "iso2c",
          destination = "iso3c"
        )
  )
```

## Combine with weird indicators
```{r}
# open weird data
DF.w <-
  readRDS(file.path(
    data_dir,
    "worlddata",
    "btpsych24a_weird.Rds"
  ))

# for simplicity, calculate average 'ride' scores for each country
DF.w <- DF.w %>%
  group_by(iso3) %>%
  summarise(
    r = mean(r, na.rm = T),
    i = mean(i, na.rm = T),
    d = mean(d, na.rm = T),
    e = mean(e, na.rm = T)
  )

# left join with DF
DF.o.w <-
  left_join(DF.o,
    DF.w,
    by = "iso3"
  )

# export
saveRDS(DF.o.w,
  file = file.path(
    data_dir,
    "worlddata",
    "DF.o.w.Rds"
  )
)

# reload
DF.o.w <- readRDS(
  file =
    file.path(
      data_dir,
      "worlddata",
      "DF.o.w.Rds"
    )
)

# prepare population data
DF.pop <-
  read.csv(
    file.path(
      data_dir,
      "worlddata",
      "population",
      "btpsych24a_worlddata_pop.csv"
    ),
    skip = 4
  )

# process population data
DF.pop <- DF.pop %>%
  filter(Country.Name != "Not classified") %>%
  rowwise() %>%
  mutate(pop = mean(X2000:X2024, na.rm = T)) %>%
  ungroup() %>%
  select(Country.Code, pop)

# prepare merge with other indicators
DF.w2 <- DF.w %>%
  group_by(iso3) %>%
  summarise(r = mean(r, na.rm = T), i = mean(i, na.rm = T), d = mean(d, na.rm = T), e = mean(e, na.rm = T))

tmp <- DF.pop %>%
  rename("iso3" = "Country.Code") %>%
  left_join(DF.w2, by = "iso3")

# export
saveRDS(tmp,
  file = file.path(
    data_dir,
    "worlddata",
    "tmp.Rds"
  )
)

tmp <- 
  readRDS(file = 
            file.path(data_dir,
                      "worlddata",
                      "tmp.Rds")
          )
```

# Create plots comparing distribution across different bands of big team science
Create general plotting function
```{r}
DataViz2 <- function(outcome,
                     label,
                     adjust = 1) {
  ggplot(
    data = DF.o.w,
    aes(x = .data[[outcome]])
  ) +

    # size.bin == '1 - 4'
    geom_density(
      data = DF.o.w %>%
        filter(size.bin == '1 - 4'),
      color = "#1b9e77",
      adjust = adjust,
      alpha = 0.4
    ) +
    
    # size.bin == '5 - 9'
    geom_density(
      data = DF.o.w %>%
        filter(size.bin == '5 - 9'),
      color = "#d95f02",
      adjust = adjust,
      alpha = 0.4
    ) +

    # author.n > 10
    geom_density(
      data = DF.o.w %>%
        filter(authors.n > 10),
      color = "#7570b3",
      adjust = adjust,
      alpha = 0.4
    ) +

    # actual distribution
    geom_density(
      data = tmp,
      aes(weight = pop),
      linetype = "dotted",
      adjust = adjust,
      alpha = 0.4
    ) +

    # fix aes
    xlab(label) +
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.title.y = element_blank()
    )
}
```

Create figures
```{r}
r.fig <-
  DataViz2(
    outcome = "r",
    label = "GDP Per Capita (Millions)",
    adjust = 5
  )

i.fig <-
  DataViz2(
    outcome = "i",
    label = "Industry-Based GDP (%)",
    adjust = 5
  )

d.fig <-
  DataViz2(
    outcome = "d",
    label = "Democracy Index",
    adjust = 5
  )

e.fig <-
  DataViz2(
    outcome = "e",
    label = "Secondary School Enrollment (% Gross)",
    adjust = 5
  )

base_plot <-
  plot_grid(
    r.fig,
    i.fig,
    d.fig,
    e.fig
  )

legend_levels <-
  c(
    "1st Author Affiliation (< 10 Co-Authors)",
    "1st Author Affiliation (10+ Co-Authors)",
    "Between-country distribution"
  )

# dummy df
legend_df <- data.frame(
  size.bin = factor(legend_levels, levels = legend_levels),
  x = c(1, 2, 3),
  y = c(1, 1, 1)
)

# color & linetype mapping (named vector ensures mapping by name)
my_colors <- c(
  "1st Author Affiliation (< 10 Co-Authors)" = "#4cc9f0",
  "1st Author Affiliation (10+ Co-Authors)" = "#7b2cbf",
  "Between-Country Distribution" = "black"
)

my_linetypes <- c(
  "1st Author Affiliation (< 10 Co-Authors)" = "solid",
  "1st Author Affiliation (10+ Co-Authors)" = "solid",
  "Between-country distribution" = "dotted"
)

# 2. Build dummy plot mapping aesthetics at top-level

# Make a custom legend
legend_labels <-
  c(
    "1st Author Affiliation (< 10 Co-Authors)",
    "1st Author Affiliation (10+ Co-Authors)",
    "Between-Country Distribution"
  )

legend_colors <-
  c(
    "#4cc9f0",
    "#7b2cbf",
    "black"
  )

# Create a dummy ggplot to extract a legend
dummy <-
  ggplot(
    data.frame(
      x = legend_labels,
      y = 1,
      col = legend_labels
    ),
    aes(x,
      y,
      color = col
    )
  ) +
  geom_point() +
  scale_color_manual(
    values = legend_colors,
    labels = legend_labels
  ) +
  guides(
    color =
      guide_legend(
        title = NULL,
        direction = "horizontal"
      )
  ) +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 12)
  )

legend <- get_legend(dummy)

# Add legends to
legend_plot <-
  plot_grid(base_plot,
    legend_grob,
    nrow = 2,
    rel_heights = c(1, 1)
  )

print(legend_plot)

final_plot <- ggdraw() +
  draw_plot(legend_plot, y = 0, height = 0.9)

print(final_plot)
```

Some analyses
```{r}
DF.o.w <-
  DF.o.w %>%
  mutate(
    group =
      case_when(
        size.bin == "1 - 4" ~ "small",
        size.bin == "5 - 9" ~ "medium",
        authors.n > 9 ~ "large"
      )
  )

WEIRD_correlations <- c(NA, NA, NA, NA)

# r
kruskal.test(r ~ group,
  data = DF.o.w
)

WEIRD_correlations[1] <- cor.test(
  x = DF.o.w$r,
  y = DF.o.w$authors.n,
  alternative = "two.sided",
  method = "spearman"
)$estimate[[1]]

# i
kruskal.test(i ~ group,
  data = DF.o.w
)

WEIRD_correlations[2] <- cor.test(
  x = DF.o.w$i,
  y = DF.o.w$authors.n,
  alternative = "two.sided",
  method = "spearman"
)$estimate[[1]]

# d
kruskal.test(d ~ group,
  data = DF.o.w
)

WEIRD_correlations[3] <- cor.test(
  x = DF.o.w$d,
  y = DF.o.w$authors.n,
  alternative = "two.sided",
  method = "spearman"
)$estimate[[1]]

# e
kruskal.test(e ~ group,
  data = DF.o.w
)

WEIRD_correlations[4] <- cor.test(
  x = DF.o.w$e,
  y = DF.o.w$authors.n,
  alternative = "two.sided",
  method = "spearman"
)$estimate[[1]]


WEIRD_matrix <- t(matrix(WEIRD_correlations))

rownames(WEIRD_matrix) <- "Team size"
colnames(WEIRD_matrix) <- c(
  "GDP per capita",
  "Industrialization",
  "Democratic ratings",
  "Educational attainment"
)


corrplot(WEIRD_matrix,
  method = "ellipse",
  addCoef.col = "#36454F", number.cex = 0.9, number.digits = 3,
  tl.cex = 0.9, tl.col = "#36454F",
  col = colorRampPalette(c("#0066CC", "white", "#ff2917"))(200),
  mar = c(1, 0, 1, 1)
)
```

# John analyses

```{r}
# For each team size, calculate the percentage of scientific papers of produced by each country

# identify top five producers of psychology
top5.p <-
  DF.o.w %>% 
  filter(!is.na(iso3)) %>% 
  group_by(iso3) %>% 
  summarise(n.papers = n()) %>%
  mutate(per.papers = n.papers / sum(n.papers)) %>% 
  slice_max(per.papers, 
            n = 5)

# compare to population
top5.w <-
  tmp %>% 
  mutate(per.pop = pop / sum(pop)) %>% 
  filter(iso3 %in% top5.p$iso3)

top5.p$per.papers %>% sum()
top5.w$per.pop %>% sum()

# proportions by small, medium, big
DF.percent <- 
  DF.o.w %>%
  
  # remove non-iso data
  filter(!is.na(iso3)) %>% 
  
  # identify
  mutate(size = 
           case_when(size.bin == "1 - 4" ~ 'small',
                     size.bin == "5 - 9" ~ 'medium',
                     authors.n > 9 ~ 'big')) %>% 
  
  # group by number of authors and country
  group_by(size, iso3) %>%
  
  # calculate the total number of papers of each team size produced by country
  summarise(n_papers = n()) %>%
  ungroup() %>%
  
  # calculate the percentage of papers produced by that country
  group_by(size) %>%
  mutate(percent = 100 * n_papers / sum(n_papers)) %>%
  ungroup() %>%
  
  # sort
  arrange(size, 
          -percent)

# isolate five
DF.percent %>% 
  group_by(size) %>% 
  slice_max(percent, n = 5) %>% 
  mutate(share = sum(percent)) %>% 
  View()


tmp %>% 
  filter(iso3 %in% c('USA', 'GRB', 'CHN', 'BRA', 'CAN')) %>% 
  .$pop %>% 
  sum() / sum(tmp$pop)

# add country socioeconomic data
DF.percent <- left_join(DF.percent, DF.w, by = "iso3")

# model the impact of team size and socioeconomic factors on scientific participation
model.percent <- lm(
  data = DF.percent,
  percent ~ authors.n * (
    scale(r) + # standardize socioeconomic indices
      scale(i) +
      scale(d) +
      scale(e))
)
summary(model.percent)

# analyze the 5 countries with the most scietific output
# list the countries in iso3 code
top_countries <- c("USA", "CHN", "GBR", "DEU", "ITA")

# filter the countries of interest and restrict analysis to teamsize <=100
DF.percent.top <- DF.percent %>%
  filter(iso3 %in% top_countries &
    authors.n <= 100) %>%
  select(authors.n, iso3, percent)


model.percent.top <- DF.percent.top %>%
  lm(data = ., percent ~ authors.n * iso3)
summary(model.percent.top)

# calculate the sum of the 5 countries' scientific participation for data visualization
DF.totals <- DF.percent.top %>%
  group_by(authors.n) %>%
  summarise(percent = sum(percent, na.rm = TRUE), .groups = "drop") %>%
  mutate(iso3 = "Combined") %>%
  select(authors.n, iso3, percent)

# add combined participation to dataframe
DF.percent.top <- rbind(DF.percent.top, DF.totals)

DF.percent.top$iso3 <- factor(DF.percent.top$iso3,
  levels = c(
    "USA", "GBR", "DEU",
    "CHN", "ITA", "Combined"
  )
)

# plot scientific participation data for the highest output countries
ggplot(DF.percent.top, aes(x = authors.n, y = percent, color = iso3, group = iso3)) +
  geom_smooth(
    method = "lm",
    fill = "lightgrey",
    alpha = 0.35
  ) +
  geom_point() +
  labs(
    x = "Number of Authors",
    y = "Scientific Participation (%)",
    title = "Scientific Output Percentage by Country and Team Size"
  ) +
  scale_color_manual(
    name = "Country",
    labels = c(
      "United States", "Great Britain",
      "Germany", "China",
      "Italy", "Combined"
    ),
    values = c(
      "USA" = "#F8766D",
      "GBR" = "#B79F00",
      "DEU" = "#7CAE00",
      "CHN" = "#00BFC4",
      "ITA" = "#619CFF",
      "Combined" = "#F564E3"
    )
  ) +
  theme_minimal()


# Load required libraries
library(ggpubr)
library(cowplot)
library(ggplot2)

# Relabel size_group
DF.o.w <- DF.o.w %>%
  mutate(size_group = factor(case_when(
    size.bin == "1 - 4" ~ "1 - 4 co-authors",
    size.bin == "5 - 9" ~ "5 - 9 co-authors",
    TRUE ~ "10+ co-authors"
  ), levels = c("10+ co-authors", "5 - 9 co-authors", "1 - 4 co-authors")))

# Variable labels
var_labels <- list(
  r = "GDP Per Capita (Millions)",
  i = "Industry-Based GDP (%)",
  d = "Democracy Index",
  e = "Secondary School Enrollment (% Gross)"
)

# Function to create annotated boxplot for each variable

make_boxplot <- function(var, label) {
  ggpubr::ggboxplot(DF.o.w,
    x = "size_group", y = var,
    fill = "size_group", color = "size_group",
    palette = "jco", width = 0.5, outlier.shape = NA
  ) +
    labs(x = "Team Size Group", y = label) +
    theme_classic() +
    theme(legend.position = "none", axis.title.x = element_text(size = 12), axis.title.y = element_text(size = 12)) +
    stat_compare_means(
      method = "wilcox.test",
      label = "p.signif",
      comparisons = as.list(combn(levels(DF.o.w$size_group), 2, simplify = FALSE))
    )
}

# Create plots
r_plot <- make_boxplot("r", var_labels$r)
i_plot <- make_boxplot("i", var_labels$i)
d_plot <- make_boxplot("d", var_labels$d)
e_plot <- make_boxplot("e", var_labels$e)

# Combine plots
plot_grid(r_plot, i_plot, d_plot, e_plot, ncol = 2, labels = c("A", "B", "C", "D"))
```

# Nicholas paper analyses
## Compare global distribution to psychology first authors
```{r}
# Run weighted KS test
ks_test(x = tmp$e, 
        y = DF.o.w$e, 
        w_x = tmp$pop)

# i
ks_test(x = tmp$i, 
        y = DF.o.w$i, 
        w_x = tmp$pop)

# r
ks_test(x = tmp$r, 
        y = DF.o.w$r, 
        w_x = tmp$pop)

# d
ks_test(x = tmp$d, 
        y = DF.o.w$d, 
        w_x = tmp$pop)
```

## Examine monotonic links between team size and ride
```{r}

# spearman corr
cor.test(x = DF.o.w$e,
         y = DF.o.w$authors.n,
         method = "spearman")

cor.test(x = DF.o.w$i,
         y = DF.o.w$authors.n,
         method = "spearman")

cor.test(x = DF.o.w$r,
         y = DF.o.w$authors.n,
         method = "spearman")

cor.test(x = DF.o.w$d,
         y = DF.o.w$authors.n,
         method = "spearman")
```

# Compare small, medium, and big team
```{r}
# e gets worse as you move to medium, but not big
ks_test(x = DF.o.w[DF.o.w$size.bin == '1 - 4', ]$e, 
        y = DF.o.w[DF.o.w$size.bin == '5 - 9', ]$e)

ks_test(x = DF.o.w[DF.o.w$size.bin == '5 - 9', ]$e, 
        y = DF.o.w[DF.o.w$authors.n > 9, ]$e)

# i
ks_test(x = DF.o.w[DF.o.w$size.bin == '1 - 4', ]$i, 
        y = DF.o.w[DF.o.w$size.bin == '5 - 9', ]$i)

ks_test(x = DF.o.w[DF.o.w$size.bin == '5 - 9', ]$i, 
        y = DF.o.w[DF.o.w$authors.n > 9, ]$i)

# r
ks_test(x = DF.o.w[DF.o.w$size.bin == '1 - 4', ]$r, 
        y = DF.o.w[DF.o.w$size.bin == '5 - 9', ]$r)

ks_test(x = DF.o.w[DF.o.w$size.bin == '5 - 9', ]$r, 
        y = DF.o.w[DF.o.w$authors.n > 9, ]$r)

# d
ks_test(x = DF.o.w[DF.o.w$size.bin == '1 - 4', ]$d, 
        y = DF.o.w[DF.o.w$size.bin == '5 - 9', ]$d)

ks_test(x = DF.o.w[DF.o.w$size.bin == '5 - 9', ]$d, 
        y = DF.o.w[DF.o.w$authors.n > 9, ]$d)
```


# Compare small team to big team
```{r}
ks_test(x = DF.o.w[DF.o.w$authors.n < 10, ]$e, 
        y = DF.o.w[DF.o.w$authors.n > 9, ]$e)

ks_test(x = DF.o.w[DF.o.w$authors.n < 10, ]$i, 
        y = DF.o.w[DF.o.w$authors.n > 9, ]$i)

ks_test(x = DF.o.w[DF.o.w$authors.n < 10, ]$r, 
        y = DF.o.w[DF.o.w$authors.n > 9, ]$r)

ks_test(x = DF.o.w[DF.o.w$authors.n < 10, ]$d, 
        y = DF.o.w[DF.o.w$authors.n > 9, ]$d)
```

