---
title: "Create map figure"
output: word_document
editor_options: 
  chunk_output_type: console
---

# Setup
```{r setup, include= F}
# specify data directory
data_dir <- Sys.getenv("data_dir")

# Corrplot package for correlation plot
library(corrplot)

# source setup file
source('setup.R')
```

# Open bibliometric data and identify locations
```{r}
# open data
DF.o <- 
  # open data
  readRDS(file.path(data_dir,
                    "openalex",
                    "btpsych24a_data_bib_processed.Rds"))

# identify country of first author via indexing
DF.o$author1.country <- 
  lapply(c(1 : nrow(DF.o)),
         function(x){
           DF.o[x, ][['author']][[1]][['institution_country_code']][[1]] %>% 
             return()
           }
  )

# Prepare for merging
DF.o <- DF.o %>% 
  mutate(iso3 = 
           author1.country %>% 
           
           # convert to character and identify NAs
           as.character() %>% 
           if_else(. == "NA",
                   NA,
                   .) %>% 
           
           # convert country code iso3 format (to facilitate later merge)
           countrycode(.,
                       origin = "iso2c", 
                       destination = "iso3c")
         )
```

## Combine with weird indicators
```{r}
# open weird data
DF.w <-
  readRDS(file.path(data_dir,
                    'worlddata',
                    'btpsych24a_weird.Rds')
          )

# for simplicity, calculate average 'ride' scores for each country
DF.w <- DF.w %>% 
  group_by(iso3) %>% 
  summarise(r = mean(r, na.rm = T),
            i = mean(i, na.rm = T),
            d = mean(d, na.rm = T),
            e = mean(e, na.rm = T))

# left join with DF
DF.o.w <- 
  left_join(DF.o,
            DF.w,
            by = "iso3")
```

# Create plots comparing population and sample distribution
Create general plotting function
```{r}
DataViz <- function(outcome,
                    label){
  ggplot(data = DF.w,
         aes(x = .data[[outcome]])) +
  
  # distribution across-countries
  geom_density(color = '#0066CC',
               adjust = 2) +
  
  # distribution
  geom_density(data = DF.o.w,
               color = '#ff2917',
               adjust = 7) +
  
  xlab(label) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())
}
```

Create figures
```{r}
r.fig <- DataViz(outcome = 'r',
                 label = 'GDP per capita (millions)')

i.fig <- DataViz(outcome = 'i',
                 label = 'Industrialization')

d.fig <- DataViz(outcome = 'd',
                 label = 'Democratic ratings')

e.fig <- DataViz(outcome = 'e',
                 label = 'Educational attainment')

# Save combined base plot
base_plot <- plot_grid(r.fig,
          i.fig,
          d.fig,
          e.fig)



# Define labels and colors
labels <- c("Countries", "Published papers")
colors <- c("Countries" = "#0066CC", "Published papers" = "#ff2917")


# Generate a dummy dataframe to create legend only
legend_df <- data.frame(
  x = 1:2,
  y = 1:2,
  group = factor(labels, levels = labels)
)

dummy_plot <- ggplot(legend_df, aes(x = x, y = y, color = group)) +
  geom_line(linewidth = 0.5) +
  scale_color_manual(values = colors, name = NULL) +  # no legend title or add if you want
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.text = element_text(size = 10),
    legend.key.size = unit(1.5, "lines"),
    legend.title = element_text(face = "bold")
  )

# Extract legend grob
legend_grob <- get_legend(dummy_plot)

# Add legends to 
legend_plot <- plot_grid(
  base_plot,
  legend_grob,
  ncol = 2,
  rel_widths = c(1, 0.2)
)

print(legend_plot)

final_plot <- ggdraw() +
  draw_label("Country and Published Papers by Socioeconomic Indices", fontface = "bold", size = 12, x = 0.5, hjust = 0.5, y = 0.95) +
  draw_plot(legend_plot, y = 0, height = 0.9)

print(final_plot)
```

# Create plots comparing distribution across different bands of big team science
Create general plotting function
```{r}
DataViz2 <- function(outcome,
                     label){
  ggplot(data = DF.o.w,
         aes(x = .data[[outcome]])) +
  
  # small team distribution
  geom_density(data = DF.o.w %>% 
                 filter(size.bin == '1 - 4'),
               color = '#F8766D',
               #adjust = 10
               ) +
    
  # medium team distribution
  geom_density(data = DF.o.w %>% 
                 filter(size.bin == '5 - 9'),
               color = '#DE8C00',
               #adjust = 10
               ) +
    
  # big team distribution
  geom_density(data = DF.o.w %>% 
                 filter(authors.n > 9),
               color = '#00BFC4',
               #adjust = 10
               ) +
    
  # actual distribution
  geom_density(data = DF.w,
               linetype = 'dotted',
               #adjust = 2
               ) +
    
  # fix aes
  xlab(label) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())
}
```

Create figures
```{r}
r.fig <- DataViz2(outcome = 'r',
                  label = 'GDP per capita (millions)')

i.fig <- DataViz2(outcome = 'i',
                  label = 'Industrialization')

d.fig <- DataViz2(outcome = 'd',
                  label = 'Democratic ratings')

e.fig <- DataViz2(outcome = 'e',
                  label = 'Educational attainment')

base_plot <- plot_grid(r.fig,
          i.fig,
          d.fig,
          e.fig)

legend_levels <- c("Small teams (1–4)", "Medium teams (5–9)", "Big teams (10+)", "Country distribution")

# dummy df
legend_df <- data.frame(
  size.bin = factor(legend_levels, levels = legend_levels),
  x = c(1,2,3,4),
  y = c(1,1,1,1)
)

# color & linetype mapping (named vector ensures mapping by name)
my_colors <- c(
  "Small teams (1–4)" = "#F8766D",
  "Medium teams (5–9)" = "#DE8C00",
  "Big teams (10+)"    = "#00BFC4",
  "Country distribution" = "black"
)

my_linetypes <- c(
  "Small teams (1–4)" = "solid",
  "Medium teams (5–9)" = "solid",
  "Big teams (10+)"    = "solid",
  "Country distribution" = "dotted"
)

# 2. Build dummy plot mapping BOTH aesthetics at top-level
dummy_plot <- ggplot(legend_df, aes(x = x, y = y, color = size.bin, linetype = size.bin)) +
  geom_line(size = 1.2) +
  # IMPORTANT: give both scales the same name (or both name = "")
  scale_color_manual(
    name = "",                      # same name for both scales => combined legend
    values = my_colors,
    breaks = legend_levels          # enforce order
  ) +
  scale_linetype_manual(
    name = "",                      # same name
    values = my_linetypes,
    breaks = legend_levels
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 1.2)),   # ensure line is visible in key
    linetype = guide_legend(override.aes = list(size = 1.2)) # won't make a second legend because names match
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",      # will be right when you extract / display
    legend.key.width = unit(1.2, "cm"),
    legend.text = element_text(size = 11)
  )

# 3. Extract legend grob
legend_grob <- get_legend(dummy_plot)

# Add legends to 
legend_plot <- plot_grid(
  base_plot,
  legend_grob,
  ncol = 2,
  rel_widths = c(1, 0.3)
)

print(legend_plot)

final_plot <- ggdraw() +
  draw_label("Comparison of Small, Medium and Big Team Papers by Socioeconomic Indices", fontface = "bold", size = 12, x = 0.5, hjust = 0.5, y = 0.95) +
  draw_plot(legend_plot, y = 0, height = 0.9)

print(final_plot)



```

Some analyses
```{r}
DF.o.w <- 
  DF.o.w %>% 
  mutate(group = 
           case_when(size.bin == '1 - 4' ~ 'small',
                     size.bin == '5 - 9' ~ 'medium',
                     authors.n > 9 ~ 'large')
         )

WEIRD_correlations <- c(NA, NA, NA, NA)

# r
kruskal.test(r ~ group,
             data = DF.o.w)

WEIRD_correlations[1] <- cor.test(x = DF.o.w$r,
                                  y = DF.o.w$authors.n,
                                  alternative = 'two.sided',
                                  method = 'spearman')$estimate[[1]]

# i
kruskal.test(i ~ group,
             data = DF.o.w)

WEIRD_correlations[2] <- cor.test(x = DF.o.w$i,
                                  y = DF.o.w$authors.n,
                                  alternative = 'two.sided',
                                  method = 'spearman')$estimate[[1]]

# d
kruskal.test(d ~ group,
             data = DF.o.w)

WEIRD_correlations[3] <- cor.test(x = DF.o.w$d,
                                  y = DF.o.w$authors.n,
                                  alternative = 'two.sided',
                                  method = 'spearman')$estimate[[1]]

# e
kruskal.test(e ~ group,
             data = DF.o.w)

WEIRD_correlations[4] <- cor.test(x = DF.o.w$e,
                                  y = DF.o.w$authors.n,
                                  alternative = 'two.sided',
                                  method = 'spearman')$estimate[[1]]


WEIRD_matrix <- t(matrix(WEIRD_correlations))

rownames(WEIRD_matrix) <- "Team size"
colnames(WEIRD_matrix) <- c("GDP per capita",
                            "Industrialization",
                            "Democratic ratings",
                            "Educational attainment")


corrplot(WEIRD_matrix, method = "ellipse",
        addCoef.col = "#36454F", number.cex = 0.9, number.digits = 3,
         tl.cex = 0.9, tl.col = "#36454F", 
         col = colorRampPalette(c("#0066CC", "white", "#ff2917"))(200),
         mar = c(1,0,1,1))


```

# John exploratory analysis

```{r}

# For each team size, calculate the percentage of scientific papers of produced by each country

#identify the list of countries in DF.o.w
country_list <- unique(DF.o.w$iso3)

DF.percent <- DF.o.w %>%
  # group by number of authors and country
  group_by(authors.n, iso3) %>%
  # calculate the total number of papers of each team size produced by country
  summarise(n_papers = n(), .groups = "drop_last") %>%
  ungroup() %>%
  # fill in missing authors.n × countries combos with n_papers = NA
  complete(authors.n, iso3 = country_list, fill = list(n_papers = 0)) %>%
  # calculate the percentage of papers produced by that country
  group_by(authors.n) %>%
  mutate(percent = 100 * n_papers / sum(n_papers)) %>%
  ungroup() %>%   
  #remove entries without country data
  filter(!is.na(iso3))

# add country socioeconomic data 
DF.percent <- left_join(DF.percent, DF.w, by = "iso3")

# model the impact of team size and socioeconomic factors on scientific participation
model.percent <- lm(data = DF.percent, 
                    percent ~ authors.n*(
                        scale(r) + # standardize socioeconomic indices
                        scale(i) + 
                        scale(d) + 
                        scale(e)) )
summary(model.percent)

# analyze the 5 countries with the most scietific output
# list the countries in iso3 code
top_countries <- c("USA", "CHN", "GBR", "DEU", "ITA")

# filter the countries of interest and restrict analysis to teamsize <=100
DF.percent.top <- DF.percent %>%
  filter(iso3 %in% top_countries &
         authors.n <=100) %>%
  select(authors.n, iso3, percent)


model.percent.top <- DF.percent.top %>%
  lm(data = ., percent ~ authors.n*iso3)
summary(model.percent.top)

# calculate the sum of the 5 countries' scientific participation for data visualization
 DF.totals <- DF.percent.top %>%
  group_by(authors.n) %>%
  summarise(percent = sum(percent, na.rm = TRUE), .groups = "drop") %>%
  mutate(iso3 = "Combined") %>%
  select(authors.n, iso3, percent)

# add combined participation to dataframe
DF.percent.top <- rbind(DF.percent.top, DF.totals)

DF.percent.top$iso3 <- factor(DF.percent.top$iso3, 
                              levels = c("USA", "GBR", "DEU", 
                                         "CHN", "ITA", "Combined"))

# plot scientific participation data for the highest output countries
ggplot(DF.percent.top, aes(x = authors.n, y = percent, color = iso3, group = iso3)) +
  geom_smooth(method = "lm",
              fill = "lightgrey", 
              alpha = 0.35) +
  geom_point() +
  labs(
    x = "Number of Authors",
    y = "Scientific Participation (%)",
    title = "Scientific Output Percentage by Country and Team Size"
  ) +
  scale_color_manual(
    name = "Country",
    labels = c("United States", "Great Britain", 
               "Germany", "China",
               "Italy", "Combined"),
    values = c("USA" = "#F8766D",
               "GBR" ="#B79F00", 
               "DEU" =  "#7CAE00",
               "CHN" = "#00BFC4", 
               "ITA" = "#619CFF", 
               "Combined" = "#F564E3"
      )
  ) +
  theme_minimal()



```
