---
title: "btpsych24a_supp.refs"
author: "Joao Takayanagi"
date: "2025-10-02"
output: html_document
---
This code compiles the numbers quoted in the supplement

# Setup
```{r setup, include= F}
# setup
source("setup.R")

# specify directory
i_am("btpsych24a_paper.refs.Rmd")

# specify data directory
data_dir <- Sys.getenv("data_dir")

# turn off scientific notation
options(scipen = 999)
```

## Growth of Big Team Psychology Across Different Time Periods 
Load data
```{r}
# Load data
DF.oa.filter <-
  readRDS(file.path(data_dir, "openalex", "btpsych24a_data_bib_processed.Rds"))

DF <- DF.oa.filter
```

Function for calculating relative growth for a given time window
```{r}
calculate_growth <- function(data, start_year, end_year) {
  # Create time window label
  time_window <- paste0(start_year + 1, "-", end_year - 1)

  # Calculate midpoint for splitting the window
  mid_year <- start_year + ((end_year - start_year) / 2)

  growth_data <- data |>
    # Filter to the specified time window
    filter(publication_year > start_year, publication_year < end_year) |>

    # Split into two equal periods
    mutate(
      pub.period = if_else(publication_year > mid_year, "later", "earlier")
    ) |>

    # Calculate counts for each size.bin and period
    group_by(size.bin, pub.period) |>
    summarise(n = n(), .groups = "drop") |>

    # Pivot to get earlier and later columns
    pivot_wider(
      names_from = 'pub.period',
      values_from = n,
      id_cols = 'size.bin'
    ) |>

    # Calculate relative growth percentage
    rowwise() |>
    mutate(
      relative_growth = if_else(
        is.na(earlier) | earlier == 0,
        NA_real_,
        (later - earlier) / earlier * 100
      ),
      relative_growth = round(relative_growth, 1)
    ) |>

    # Add time window column and select final columns
    mutate(time_window = time_window) |>
    select(
      time_window,
      size_bin = size.bin,
      relative_growth_pct = relative_growth
    )

  return(growth_data)
}
```

### Sensitivity analysis: three periods
```{r}
# Define the three time windows
time_windows <- list(
  list(start = 2010, end = 2021), # 2011-2020
  list(start = 2005, end = 2016), # 2006-2015
  list(start = 2000, end = 2011) # 2001-2010
)

# Calculate growth for all time windows
growth_results <- map_dfr(
  time_windows,
  ~ calculate_growth(DF, .x$start, .x$end)
)

# Reorder columns for final table
sensitivity_table <- growth_results |>
  select(
    `Time Window` = time_window,
    `Size Bin` = size_bin,
    `Relative Growth (%)` = relative_growth_pct
  )

# view table
# view(sensitivity_table)
```

### Sensitivity analysis: three windows
```{r}
# Define the three time windows with varying widths
width_windows <- list(
  list(start = 2016, end = 2021, width = "4 year"), # 2017-2020
  list(start = 2010, end = 2021, width = "10 year"), # 2011-2020
  list(start = 2000, end = 2021, width = "20 year") # 2001-2020
)

# Calculate growth for all window widths
width_results <- map_dfr(
  width_windows,
  ~ calculate_growth(DF, .x$start, .x$end)
)

# Reorder columns for final table
width_sensitivity_table <- width_results |>
  select(
    `Time Window` = time_window,
    `Size Bin` = size_bin,
    `Relative Growth (%)` = relative_growth_pct
  )

# view table
# view(width_sensitivity_table)
```

## Alternative Statistical Treatments of Mentions 
### Whether record received any attention
In OpenAlex
```{r}
oa.sens <- DF.oa.filter %>%
  # examine whether record received any attention
  mutate(
    scholar.att = if_else(cited_by_count > 0, 1, 0)
  ) %>%
  # group by size bin
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  # calculate relative summary statistics
  summarise(
    mn.yearly_scholar = mean(citation.p.y) %>%
      round(2),
    md.yearly_scholar = median(citation.p.y) %>%
      round(2),
    mn_scholar = mean(cited_by_count) %>%
      round(2),
    md_scholar = median(cited_by_count) %>%
      round(2),
    att_scholar = mean(scholar.att) %>%
      round(2)
  )
```

In Altmetric
```{r}
# open data
DF.alt <- readRDS(
  file.path(
    data_dir,
    "altmetric",
    "btpsych24a_altmetric_processed.Rds"
  )
)

# process data
alt.sens <- DF.alt %>%
  # examine whether record received any attention
  mutate(
    policy.att = if_else(Policy.mentions > 0, 1, 0),
    news.att = if_else(news > 0, 1, 0),
    gen.att = if_else(gen > 0, 1, 0)
  ) %>%
  # group by size bin
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  # calculate relative summary statistics
  summarise(
    mn.yearly_news = mean(news.p.y) %>%
      round(2),
    md.yearly_news = median(news.p.y) %>%
      round(2),
    mn_news = mean(news) %>%
      round(2),
    md_news = median(news) %>%
      round(2),
    att_news = mean(news.att) %>%
      round(2),
    mn.yearly_social = mean(gen.p.y) %>%
      round(2),
    md.yearly_social = median(gen.p.y) %>%
      round(2),
    mn_social = mean(gen) %>%
      round(2),
    md_social = median(gen) %>%
      round(2),
    att_social = mean(gen.att) %>%
      round(2),
    mn.yearly_policy = mean(policy.p.y) %>%
      round(2),
    md.yearly_policy = median(policy.p.y) %>%
      round(2),
    mn_policy = mean(Policy.mentions) %>%
      round(2),
    md_policy = median(Policy.mentions) %>%
      round(2),
    att_policy = mean(policy.att) %>%
      round(2)
  )

# Join reults
sens <-
  full_join(oa.sens, alt.sens, by = "size.bin")
```

### Medians
```{r}
## Overall, median
DF.alt %>%
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  summarise(
    med.policy = median(Policy.mentions),
    med.news = median(news),
    med.gen = median(gen)
  )

## Yearly, median
DF.alt %>%
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  summarise(
    med.policy = median(policy.p.y),
    med.news = median(news.p.y),
    med.gen = median(gen.p.y)
  )

## Likelihood of being mentioned
DF.alt %>%
  mutate(
    Policy.mentions = if_else(Policy.mentions > 0, 1, 0),
    news = if_else(news > 0, 1, 0),
    gen = if_else(gen > 0, 1, 0)
  ) %>%
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  summarise(
    prop.policy = mean(Policy.mentions),
    prop.news = mean(news),
    prop.gen = mean(gen)
  )
```

### Links with prestige of the journal
```{r}

# Test correlation between team size and journal h-index
cor.test(
  x = DF$authors.n,
  y = DF$h_index,
  alternative = 'greater',
  method = 'spearman'
)
```

## Other Potential Units of Analysis 

```{r}
# Load dataset with country and institution data
DF.o.w <- readRDS(
  file.path(
    data_dir,
    "worlddata",
    "DF.o.w.Rds"
  )
)

# fix minor coding issue
DF.o.w[DF.o.w$unique_institutions == 0, ]$unique_institutions <- NA
DF.o.w[DF.o.w$unique_country_codes == 0, ]$unique_country_codes <- NA

# calculate mean and median number of institutions and countries involved in the work, by size bin
DF.o.w %>%
  group_by(size.bin) %>%
  summarise(
    mean.i = mean(unique_institutions, na.rm = T),
    median.i = median(unique_institutions, na.rm = T),
    mean.c = mean(unique_country_codes, na.rm = T),
    median.c = median(unique_country_codes, na.rm = T)
  )

# calculate correlations between potential units of analysis
## author number and number of unique institutions involved in the work
cor.test(
  DF.o.w$authors.n,
  DF.o.w$unique_institutions,
  method = "spearman"
)

## author number and number of unique institutions involved in the work
cor.test(
  DF.o.w$authors.n,
  DF.o.w$unique_country_codes,
  method = "spearman"
)

# Calculate the correlation between different units of analysis and variables of interest
## correlation between number of countries involved and scholarly citations
cor.test(
  DF.o.w$unique_country_codes,
  DF.o.w$citation.p.y,
  method = "spearman"
)
```

Look at Altmetric data
```{r}
# Join Altmetric data with institution and country data
DF.alt.w <-
  left_join(DF.alt, DF.o.w, by = "id")

## correlation between number of countries involved and news mentions
cor.test(
  DF.alt.w$unique_country_codes,
  DF.alt.w$news.p.y,
  method = "spearman"
)

## correlation between number of countries involved and social media mentions
cor.test(
  DF.alt.w$unique_country_codes,
  DF.alt.w$gen.p.y,
  method = "spearman"
)

## correlation between number of countries involved and policy mentions
cor.test(
  DF.alt.w$unique_country_codes,
  DF.alt.w$policy.p.y,
  method = "spearman"
)
```

## Division of Labor in Ultra-Large Studies
```{r}
# Number of records included in the author contribution coding analysis
DF.c <-
  # open data
  read.csv(
    file.path(
      data_dir,
      "contributorship",
      "btpsych24a_cont_coding.csv"
    ),
    na.string = "N/A"
  )

nrow(DF.c)

# number of records actually coded
DF.c <- DF.c %>%
  filter(
    # remove instances without contributorship statements
    !is.na(contributorship.statement),
    contributorship.statement != "",

    # remove instances where coders recommended removal
    remove.revise != 1,

    # remove instances where there wasn't anything to code
    !(is.na(conception) &
      is.na(execution) &
      is.na(analysis) &
      is.na(draft.writing) &
      is.na(review.writing) &
      is.na(funding))
  ) %>%
  # select relevant variables
  select(id, contributorship.statement:funding)

nrow(DF.c)
```

# Check Contributorship coding reliability
```{r}
# recoded records from second rater
DF.rater2 <-
  read.csv(
    file.path(data_dir, 'contributorship', 'btpsych24a_cont_coding_rel.csv'),
    na.string = "N/A"
  ) %>%

  # select relevant variables
  select(id, conception:draft.writing, Coder)

# original coded records
DF.rater1 <-
  read.csv(
    file.path(data_dir, 'contributorship', 'btpsych24a_cont_coding.csv'),
    na.string = "N/A"
  ) %>%

  # select relevant variables
  select(id, conception:draft.writing, Coder) %>%

  # semi join (focusing only on records that have been recoded by the second reviewer)
  semi_join(y = DF.rater2, by = 'id')

# bind files
DF.rel <- rbind(DF.rater2, DF.rater1) %>%

  # pivot data for processing
  pivot_longer(conception:draft.writing) %>%

  # count number of authors who have contributed to each category
  mutate(
    # if N/A leave blank, otherwise count number of commas and add 1
    n = if_else(
      is.na(value),
      NA,

      # otherwise, count number of commas and add 1
      str_count(value, ",") + 1
    )
  ) %>%

  select(-value) %>%

  # pivot
  pivot_wider(names_from = Coder, values_from = n) %>%
  arrange(G)
```

# Examine reliability
```{r}
ICC(
  DF.rel %>%
    select(G, J)
)
```


## Averaging across all big team psychology efforts...
```{r}
# Load author contribution processed data
DF.c <- readRDS(
  file.path(
    data_dir,
    "contributorship",
    "btpsych24a_cont_processed.Rds"
  )
)

# Overall proportion of collaborators that contribute to the different tasks in a project by task category
DF.c %>%
  mutate(prop.n = prop.n * 100) %>%
  group_by(category) %>%
  summarise(
    m = mean(prop.n, na.rm = T),
    sd = sd(prop.n, na.rm = T)
  )

# Proportion of papers in which most authors contribute to a single/dominant task, organized by task category
DF.c %>%
  filter(dominant == "yes") %>%
  group_by(category.main) %>%
  summarise(prop = (n() / nrow(DF.c)) * 100) %>%
  mutate(other = 100 - sum(prop))

# Range of the proportion of authors that contributed to the dominant task in conception-focused papers
DF.c %>%
  filter(
    category.main == "conceptualization focused (14%)",
    dominant == "yes",
    category == "conception"
  ) %>%
  summarise(
    min = min(prop.n, na.rm = T) * 100,
    max = max(prop.n, na.rm = T) * 100
  )
```



