---
title: "Create map figure"
author: "Anonymized for peer review (NC)"
reviewer: "Anonymized for peer review (JT)"
output: word_document
editor_options: 
  chunk_output_type: console
---
NICK'S BIRTHDAY IS TODAY! JULY 30

This code creates a figure that (a) illustrates the geographical distribution of big team leadership and participation rates in psychology, and (b) examines links between those rates and indicators of economic wealth (GDP)

# Setup
```{r setup, include= F}
# specify data directory
data_dir <- Sys.getenv("data_dir")

# source setup file
source('setup.R')
```

# open data files
```{r}
DF.d <- 
  read_csv(file.path(data_dir,
                     "worlddata",
                     "demo",
                     "btpsych24a_demo.csv"),
           show_col_types = F
          )

DF.e <- 
  read_csv(file.path(data_dir,
                     "worlddata",
                     "edu",
                     "btpsych24a_edu.csv"),
           show_col_types = F,
           skip = 4
          )

DF.r <- 
  read_csv(file.path(data_dir,
                     "worlddata",
                     "gdp.pc",
                     "btpsych24a_worlddata_gdp.pc.csv"),
           show_col_types = F,
           skip = 4
          )

DF.i <- 
  read.csv(file.path(data_dir,
                     "worlddata",
                     "gdp.sector",
                     "btpsych24a_gdp.sector.csv"),
           na.strings = c("..")
           )
```

# prep for merging
Consistent variable naming and format

## DF.r
```{r}
DF.r <- DF.r %>% 
  # rename variables
  rename(iso3 = 'Country Code') %>% 
  
  # select necessary variables
  select(iso3, 
         `2000` : `2023`) %>% 
  
  # pivot longer
  pivot_longer(cols = c(`2000` : `2023`),
               names_to = 'year',
               values_to = 'r', # name of dataset
               )
```

## DF.i
```{r}
DF.i <- DF.i %>% 
  # rename variables
  rename(iso3 = 'Country.Code') %>% 
  
  # select necessary variables
  select(iso3, 
         `X2000..YR2000.` : `X2024..YR2024.`) %>% 
  
  # pivot longer
  pivot_longer(cols = c(`X2000..YR2000.` : `X2024..YR2024.`),
               names_to = 'year',
               values_to = 'i', # name of dataset
               ) %>% 
   
  # fix structure of year variable
  mutate(year = 
           substr(x = year,
                  start = 2,
                  stop = 5))
```

## DF.e (to-do)
```{r}
DF.e <- DF.e %>%
  # rename variables
  rename(iso3 = 'Country Code') %>%
  
  # select necessary variables
  select(iso3,
         '2000' : '2005') %>%
  
  # pivot longer
  pivot_longer(cols = c('2000' : '2005'),
               names_to = 'year',
               values_to = 'e',
               )
  
```

## DF.d (to-do)
```{r}
DF.d <- DF.d %>%
  # rename variables
  rename(iso3 = 'Code',
        year = 'Year',
        d = 'Democracy index') %>%

  # select necessary variables
  select(iso3,
         year,
         d) %>%
  
  # fix structure of year variable
  mutate(year = as.character(year))

```

# join (i.e., merge)
Note: dataframes need the same variable names, classes, and structures

## DF.r and DF.i = DF.ri
```{r}
DF.ri <-
  full_join(x = DF.r,
            y = DF.i,
            by = c('iso3', 
                   'year'))
```

## DF.ri and DF.d = DF.rid
```{r}
DF.rid <-
  full_join(x = DF.ri,
            y = DF.d,
            by = c('iso3',
                   'year' ))
```

## DF.rid and DF.e = DF.ride
```{r}
DF.ride <-
  full_join(x = DF.rid,
            y = DF.e,
            by = c('iso3',
                   'year'))
```

# Export as .Rds
```{r}
DF.ride %>% 
  saveRDS(file = file.path(data_dir,
                           "worlddata",
                           'btpsych24a_weird.Rds')
          )

# delete vestigial
rm(DF.d, DF.e, DF.i, #DF.r, 
   DF.ri, DF.rid)
```

# Open geocoded data
```{r}
f.author <- 
  DF.ride 
```

