---
title: "Create impact figure"
author: "Anonymized for peer review (NC)"
reviewer: "Anonymized for peer review (JT)"
output: word_document
editor_options: 
  chunk_output_type: console
---
This code prepares a figure that illustrates links between team size and multiple measures of impact

# Setup
```{r setup, include= F}
# load libaries
library('tidyverse')
library('cowplot')
library('here')
library('readxl')
library('ggtext')

# turn off scientific notation
options(scipen = 999)

# set plotting theme
theme_set(theme_classic()) 

# specify directory
i_am('btpsych24a_impact.Rmd')
```

# Open and process data

## OpenAlex
```{r open.alex, include = F}
# open data
DF <- 
  readRDS(here("data",
               "openalex",
               "btpsych24a_data_bib_processed.Rds"))
```

## Altmetric
Important note: Altmetric data is proprietary. However, we have provided scrambled copies of the data so that computational reproducibility can be assessed.

Necessary edits to the code to check computational reproducibility are noted in comments.
```{r altmetric, include = F}
# compile information about unique identifiers
DF.ident <- 
  
  # pull from doi sheets
  rbind(read.csv(here("data",
                      "altmetric",
                      "btpsych24a_bib_doi1.csv")
                 ),
        read.csv(here("data",
                      "altmetric",
                      "btpsych24a_bib_doi2.csv")
                 )
        ) %>% 
  
  # format so that the doi is similar to the Altmetric output
  rename(DOI = doi) %>% 
  mutate(DOI = str_replace(string = DOI,
                           pattern = "https://doi.org/",
                           replacement = "")) %>% 
  
  # select the identifiers that will link all the datasets
  select(openalex, DOI) %>% 
  
  # remove duplicates
  distinct(DOI, 
           .keep_all = T)

# combine two Altmetric datasets
DF.alt <-  
  rbind(
    # change to btpsych24a_altmetric1_scramble.csv if you were not provided access to raw, proprietary data
    read.csv(here("data",
                  "altmetric",
                  "btpsych24a_altmetric1_raw.csv")
             ),
    
    # change to btpsych24a_altmetric2_scramble.csv if you were not provided access to raw, proprietary data
    read.csv(here("data",
                  "altmetric",
                  "btpsych24a_altmetric2_raw.csv")
             )
        )

# connect the Altmetric dataset and identifiers 
DF.alt <-
  inner_join(x = DF.alt,
             y = DF.ident,
             by = "DOI") %>% 
  
  # give consistent naming to OpenAlex identifier
  rename(id = openalex)

# connect the identified Altmetric dataset with the OpenAlex data
DF.alt <-
  inner_join(x = DF.alt,  
             y = DF %>% 
              select(id, size.bin, publication_year, authors.n),
             by = "id") %>% 
  
  # calculate news and general population mentions
  mutate(news = News.mentions + Blog.mentions,
         gen = X.mentions + Facebook.mentions) 

# calculate yearly mentions
DF.alt <- DF.alt %>% 
  mutate(policy.p.y = Policy.mentions / (2025 - publication_year),
         news.p.y = news / (2025 - publication_year),
         gen.p.y = gen / (2025 - publication_year))

# delete vestigial
rm(DF.ident)
```

Export cleaned Altmetric data
```{r}
# select necessary variables
DF.alt <-
  DF.alt %>% 
  select(id : gen.p.y, Policy.mentions)

saveRDS(DF.alt,
        here('data',
             'altmetric',
             'btpsych24a_altmetric_processed.Rds')
        )
```

# Prep figure data
```{r}
# calculate mean yearly citation data for each size bin
cit.DF <- DF %>% 
  group_by(size.bin) %>% 
  summarise(
    citation.m = mean(citation.p.y),
    citation.sd = sd(citation.p.y),
    citation.se = sd(citation.p.y) / sqrt(n()),
    citation.n = n()
    )

# do the same for policy, news, and general media mentions
other.DF <- DF.alt %>% 
  group_by(size.bin) %>% 
  summarise(policy.m = mean(policy.p.y),
            policy.sd = sd(policy.p.y),
            policy.se = sd(policy.p.y) / sqrt(n()),
            
            news.m = mean(news.p.y),
            news.sd = sd(news.p.y),
            news.se = sd(news.p.y) / sqrt(n()),
            
            social.m = mean(gen.p.y),
            social.sd = sd(gen.p.y),
            social.se = sd(gen.p.y) / sqrt(n())
            )

# combine and prep for visualization
viz.DF <- 
  # join the summary dataframes
  full_join(cit.DF, other.DF,
            by = "size.bin") %>% 
  
  # remove blank columns  
  filter(!is.na(size.bin)) %>% 
  
  # restructure
  pivot_longer(cols = citation.m : 
                 social.se) %>%
  separate(col = name,
           into = c('outcome', 'stat'),
           sep = "\\.") %>% 
  pivot_wider(names_from = 'stat',
              values_from = 'value') %>% 
  
  # refactor
  mutate(outcome = factor(outcome,
                          levels = c('citation',
                                     'news', 
                                     'social', 
                                     'policy'), 
                          labels = c("scholarly articles", 
                                     "news", 
                                     "social media", 
                                     "policy documents")
                          )
  )

# delete vestigial
rm(cit.DF, other.DF)
```

Back up viz.DF
```{r}
viz.DF %>% 
  saveRDS(here('figures',
               'impact.viz.DF.Rds'))
```

Optionally reload viz.DF
```{r eval = F}
viz.DF <- readRDS(here('figures',
                       'impact.viz.DF.Rds')
                  )
```

# Create figure
Base plot: Bar chart showing links between team size and four types of impact
```{r}
impact.fig <- 
  ggplot(viz.DF,
       aes(x = size.bin, 
           y = m, 
           fill = size.bin)) +
  
  # facet by outcome
  facet_wrap(vars(outcome),
             nrow = 2,
             ncol = 2,
             scales = "free_y") +
  
  # barchart
  geom_bar(stat = 'identity',
           alpha = .6) +
  
  # error bars
  geom_errorbar(aes(ymin = m - se,
                    ymax = m + se),
                width = .5,
                colour = "grey50") +
  
  # improve aesthetics
  labs(x = 'team size',
       y = 'average yearly mentions') +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90,
                                   vjust = .5,
                                   hjust = 1))
```

# Export raw figure data
```{r}
# join summary data with original continuous data
viz.DF2 <- 
  inner_join(DF %>%  # JT: changed to inner join. Was previously left join
              select(id, size.bin, authors.n, citation.p.y),
            DF.alt %>% 
              select(id, policy.p.y,
                     news.p.y,
                     gen.p.y),
            by = "id") 

# export
viz.DF2 %>% 
  saveRDS(here('figures',
               'impact.viz.DF.raw.Rds'))
```

# Make continuous graphs
Optionally re-load dataset
```{r eval = F}
viz.DF2 <- 
  readRDS(here('figures',
               'impact.viz.DF.raw.Rds'))
```

```{r}
panel.figs <-
  lapply(c('citation.p.y', 'news.p.y', 
           'gen.p.y', 'policy.p.y'),
         function(outcome){
           ggplot(viz.DF2 %>% 
                    # filter out these instances for consistent scaling
                    filter(authors.n < 100),
                  aes(x = authors.n,
                      y = .data[[outcome]])) +
           
           # model linear relationship
           geom_smooth(method = "lm", 
                       formula = y ~ x, 
                       color = "grey50",
                       linetype = 'dashed',
                       linewidth = .75) +
           
           # model quadratic relationsnip
           geom_smooth(method = "lm", 
                       formula = y ~ x + I(x^2),
                       color = "grey50",
                       linewidth =  .75) +
             
           # improve aesthetics
           theme(axis.title.y = element_blank(),
                 axis.title.x = element_blank()) +
           scale_y_continuous(breaks = scales::breaks_extended(n = 4)) +
           scale_x_continuous(breaks = c(1, 50, 100))
           }
         )
```

Combine base and facet plots
```{r}
ggdraw() +
  draw_plot(impact.fig) +
  draw_plot(panel.figs[[1]], 
            x = .1,
            y = .75,
            width = .18,
            height = .18) +
  draw_plot(panel.figs[[2]], 
            x = .575,
            y = .75,
            width = .18,
            height = .18) +
  draw_plot(panel.figs[[3]], 
            x = .1,
            y = .3,
            width = .18,
            height = .18) +
  draw_plot(panel.figs[[4]], 
            x = .575,
            y = .3,
            width = .18,
            height = .18)
```
