---
title: "Create map figure"
author: "Anonymized for peer review (NC)"
reviewer: "Anonymized for peer review (JT)"
output: word_document
editor_options: 
  chunk_output_type: console
---

This code creates a figure that (a) illustrates the geographical distribution of big team leadership and participation rates in psychology, and (b) examines links between those rates and indicators of economic wealth (GDP)

# Setup
```{r setup, include= F}
# specify data directory
data_dir <- Sys.getenv("data_dir")

# source setup file
source('setup.R')
```

# Prepare OpenAlex data
## Open and isolate OpenAlex authorship data
```{r open.alex, include = F}
DF.author <- 
  # open data
  readRDS(file.path(data_dir,
                    "openalex",
                    "btpsych24a_data_bib_processed.Rds")) %>% 
  
  # isolate big team science records
  filter(size.bin != "1 - 4",
         size.bin != "5 - 9")

# focus on author lists
DF.author <- DF.author$author
```

## Geocode records
```{r}
DF.geo <- 
  # isolate affiliation information
  lapply(DF.author,
         function(sublist)
           c(sublist['institution_display_name'],
             sublist['institution_country_code'])
         ) %>% 
  
  # select only instances with full affiliation information
  ## changed to *any insatance* on Jul 28 2025 following reviewer recommendation
  ### this drops number of records from 49919 to 41107
  #Filter(function(sublist) all(!is.na(unlist(sublist))), .) %>% 
  Filter(function(sublist) any(!is.na(unlist(sublist))), .) %>% 
  
  # geocode using institution name and country
  lapply(.,
         function(sublist)
           as_tibble(sublist) %>% 
           mutate(address = paste0(institution_display_name,
                                   ", ",
                                   institution_country_code)
                  ) %>% 
           tidygeocoder::geocode(.,
                                 address = address,
                                 method = 'arcgis', 
                                 lat = latitude, 
                                 long = longitude
                                 )
         )

# save geocoded list
saveRDS(DF.geo,
        file.path(data_dir,
                  'location',
                  'btpsych24a_geocode.Rds')
        )
```

## Identify leaders and supporters
```{r}
# reload data (if needed)
DF.geo <- 
  readRDS(here('data',
               'location',
               'btpsych24a_geocode.Rds')
          )

DF.geo <-
  
  # separate leader and supporter location information
  lapply(DF.geo,
         function(sublist)
           sublist %>% 
           mutate(
             leader_coun = .[['institution_country_code']][1],
             leader_lat = .[["latitude"]][1],
             leader_lon = .[["longitude"]][1],
             supporter_coun = .[['institution_country_code']],
             supporter_lat = .[["latitude"]],
             supporter_lon = .[["longitude"]]
           )
         )
```

## Identify regions
```{r}
DF.geo <- 
  lapply(DF.geo,
         function(sublist)
           sublist %>% 
           mutate(
             # leader region
             leader_region = countrycode(.[['leader_coun']], 
                                         origin = 'iso2c', 
                                         destination = 'un.regionsub.name'),
             # supporter region
             supporter_region = countrycode(.[['supporter_coun']], 
                                            origin = 'iso2c', 
                                            destination = 'un.regionsub.name')
           )
         )
```
