---
title: "btpsych24a_supp.refs"
author: "Joao Takayanagi"
date: "2025-10-02"
output: html_document
---
This code compiles the numbers quoted in the supplement

# Setup
```{r setup, include= F}
# setup
source("setup.R")

# specify directory
i_am("btpsych24a_paper.refs.Rmd")

# specify data directory
data_dir <- Sys.getenv("data_dir")

# turn off scientific notation
options(scipen = 999)
```

# SUPPLEMENTAL ANALYSIS


## Growth of Big Team Psychology Across Different Time Periods 

```{r}
# Load data
DF.oa.filter <-
  readRDS(file.path(data_dir, "openalex", "btpsych24a_data_bib_processed.Rds"))

DF <- DF.oa.filter
```


```{r}
# Calculate relative growth for a given time window

## Create function
calculate_growth <- function(data, start_year, end_year) {
  # Create time window label
  time_window <- paste0(start_year + 1, "-", end_year - 1)

  # Calculate midpoint for splitting the window
  mid_year <- start_year + ((end_year - start_year) / 2)

  growth_data <- data |>
    # Filter to the specified time window
    filter(publication_year > start_year, publication_year < end_year) |>

    # Split into two equal periods
    mutate(
      pub.period = if_else(publication_year > mid_year, "later", "earlier")
    ) |>

    # Calculate counts for each size.bin and period
    group_by(size.bin, pub.period) |>
    summarise(n = n(), .groups = "drop") |>

    # Pivot to get earlier and later columns
    pivot_wider(
      names_from = 'pub.period',
      values_from = n,
      id_cols = 'size.bin'
    ) |>

    # Calculate relative growth percentage
    rowwise() |>
    mutate(
      relative_growth = if_else(
        is.na(earlier) | earlier == 0,
        NA_real_,
        (later - earlier) / earlier * 100
      ),
      relative_growth = round(relative_growth, 1)
    ) |>

    # Add time window column and select final columns
    mutate(time_window = time_window) |>
    select(
      time_window,
      size_bin = size.bin,
      relative_growth_pct = relative_growth
    )

  return(growth_data)
}
```

## Sensitivity analysis: three periods
```{r}
# Define the three time windows
time_windows <- list(
  list(start = 2010, end = 2021), # 2011-2020
  list(start = 2005, end = 2016), # 2006-2015
  list(start = 2000, end = 2011) # 2001-2010
)

# Calculate growth for all time windows
growth_results <- map_dfr(
  time_windows,
  ~ calculate_growth(DF, .x$start, .x$end)
)

# Reorder columns for final table
sensitivity_table <- growth_results |>
  select(
    `Time Window` = time_window,
    `Size Bin` = size_bin,
    `Relative Growth (%)` = relative_growth_pct
  )

# view table
view(sensitivity_table)
```


```{r}
## Sensitivity analysis: three windows

# Define the three time windows with varying widths
width_windows <- list(
  list(start = 2016, end = 2021, width = "4 year"), # 2017-2020
  list(start = 2010, end = 2021, width = "10 year"), # 2011-2020
  list(start = 2000, end = 2021, width = "20 year") # 2001-2020
)

# Calculate growth for all window widths
width_results <- map_dfr(
  width_windows,
  ~ calculate_growth(DF, .x$start, .x$end)
)

# Reorder columns for final table
width_sensitivity_table <- width_results |>
  select(
    `Time Window` = time_window,
    `Size Bin` = size_bin,
    `Relative Growth (%)` = relative_growth_pct
  )

# view table
view(width_sensitivity_table)
```

## Alternative Statistical Treatments of Mentions 

```{r}
### Alternative statistical operationalization of impact (OpenAlex)

oa.sens <- DF.oa.filter %>%
  # examine whether record received any attention
  mutate(
    scholar.att = if_else(cited_by_count > 0, 1, 0)
  ) %>%
  # group by size bin
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  # calculate relative summary statistics
  summarise(
    mn.yearly_scholar = mean(citation.p.y) %>%
      round(2),
    md.yearly_scholar = median(citation.p.y) %>%
      round(2),
    mn_scholar = mean(cited_by_count) %>%
      round(2),
    md_scholar = median(cited_by_count) %>%
      round(2),
    att_scholar = mean(scholar.att) %>%
      round(2)
  )
```

```{r}
### Alternative statistical operationalization of impact Altmetric

# open Altmetric data
DF.alt <- readRDS(
  file.path(
    data_dir,
    "altmetric",
    "btpsych24a_altmetric_processed.Rds"
  )
)
```


```{r}
# sensitivity analyses
alt.sens <- DF.alt %>%
  # examine whether record received any attention
  mutate(
    policy.att = if_else(Policy.mentions > 0, 1, 0),
    news.att = if_else(news > 0, 1, 0),
    gen.att = if_else(gen > 0, 1, 0)
  ) %>%
  # group by size bin
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  # calculate relative summary statistics
  summarise(
    mn.yearly_news = mean(news.p.y) %>%
      round(2),
    md.yearly_news = median(news.p.y) %>%
      round(2),
    mn_news = mean(news) %>%
      round(2),
    md_news = median(news) %>%
      round(2),
    att_news = mean(news.att) %>%
      round(2),
    mn.yearly_social = mean(gen.p.y) %>%
      round(2),
    md.yearly_social = median(gen.p.y) %>%
      round(2),
    mn_social = mean(gen) %>%
      round(2),
    md_social = median(gen) %>%
      round(2),
    att_social = mean(gen.att) %>%
      round(2),
    mn.yearly_policy = mean(policy.p.y) %>%
      round(2),
    md.yearly_policy = median(policy.p.y) %>%
      round(2),
    mn_policy = mean(Policy.mentions) %>%
      round(2),
    md_policy = median(Policy.mentions) %>%
      round(2),
    att_policy = mean(policy.att) %>%
      round(2)
  )

sens <-
  full_join(oa.sens, alt.sens, by = "size.bin")

view(sens)
```

```{r}
## Overall, median
DF.alt %>%
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  summarise(
    med.policy = median(Policy.mentions),
    med.news = median(news),
    med.gen = median(gen)
  )

## Yearly, median
DF.alt %>%
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  summarise(
    med.policy = median(policy.p.y),
    med.news = median(news.p.y),
    med.gen = median(gen.p.y)
  )

## Likelihood of being mentioned
DF.alt %>%
  mutate(
    Policy.mentions = if_else(Policy.mentions > 0, 1, 0),
    news = if_else(news > 0, 1, 0),
    gen = if_else(gen > 0, 1, 0)
  ) %>%
  filter(!is.na(size.bin)) %>%
  group_by(size.bin) %>%
  summarise(
    prop.policy = mean(Policy.mentions),
    prop.news = mean(news),
    prop.gen = mean(gen)
  )
```


```{r}
## Monotonic relationship between team size and prestige of the journal they publish in

# Test correlation between team size and journal h-index
cor.test(x = DF$authors.n, 
         y = DF$h_index,
         alternative = 'greater',
         method = 'spearman')
```

## Other Potential Units of Analysis 

```{r}


```

## Division of Labor in Ultra-Large Studies
```{r}
# Number of records included in the author contribution coding analysis
DF.c <-
  # open data
  read.csv(
    file.path(
      data_dir,
      "contributorship",
      "btpsych24a_cont_coding.csv"
    ),
    na.string = "N/A"
  )

nrow(DF.c)

# number of records actually coded
DF.c <- DF.c %>%
  filter(
    # remove instances without contributorship statements
    !is.na(contributorship.statement),
    contributorship.statement != "",

    # remove instances where coders recommended removal
    remove.revise != 1,

    # remove instances where there wasn't anything to code
    !(is.na(conception) &
      is.na(execution) &
      is.na(analysis) &
      is.na(draft.writing) &
      is.na(review.writing) &
      is.na(funding))
  ) %>%
  # select relevant variables
  select(id, contributorship.statement:funding)

nrow(DF.c)
```

## Big teams are mobilized (supplemental analysis)
```{r}

# Load author contribution processed data
DF.c <- readRDS(
  file.path(
    data_dir,
    "contributorship",
    "btpsych24a_cont_processed.Rds"
  )
)

# Overall proportion of collaborators that contribute to the different tasks in a project by task category
DF.c %>%
  mutate(prop.n = prop.n * 100) %>%
  group_by(category) %>%
  summarise(
    m = mean(prop.n, na.rm = T),
    sd = sd(prop.n, na.rm = T)
  )

# Proportion of papers in which most authors contribute to a single/dominant task, organized by task category
DF.c %>%
  filter(dominant == "yes") %>%
  group_by(category.main) %>%
  summarise(prop = (n() / nrow(DF.c)) * 100) %>%
  mutate(other = 100 - sum(prop))

# Range of the proportion of authors that contributed to the dominant task in conception-focused papers
DF.c %>%
  filter(
    category.main == "conceptualization focused (14%)",
    dominant == "yes",
    category == "conception"
  ) %>%
  summarise(
    min = min(prop.n, na.rm = T) * 100,
    max = max(prop.n, na.rm = T) * 100
  )
```



## JT Not sure where to place this


## Sensitivity analyses
### Linear and non-linear regression with massive team size outliers included
```{r}

# Download impact dataframe

DF.impact <-
  readRDS(here(
    "figures",
    "impact.viz.DF.raw.Rds"
  ))

# number of papers with over 100 articles
DF.impact %>%
  filter(authors.n > 99) %>%
  nrow()

DF.impact$authors.n %>% max()

DF.impact.full <- DF.impact %>%
  rename(
    "scholarly articles" = "citation.p.y",
    "news" = "news.p.y",
    "social media" = "gen.p.y",
    "policy documents" = "policy.p.y"
  ) %>%
  pivot_longer("scholarly articles":"social media")

ggplot(
  data = DF.impact.full,
  aes(
    x = authors.n,
    y = value
  )
) +
  facet_wrap(~name, scales = "free") +
  # model linear relationship
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    color = "grey50",
    linetype = "dashed",
    linewidth = .75
  ) +

  # model quadratic relationsnip
  geom_smooth(
    method = "lm",
    formula = y ~ x + I(x^2),
    color = "grey50",
    linewidth = .75
  )
```

