---
title: "OECD and UNESCO Data Processing"
format: html
editor_options: 
  chunk_output_type: console
---

# Setup

```{r setup, include=FALSE}
# Specify data directory
data_dir <- Sys.getenv("data_dir")

# Add source setup file
source("setup.R")
```

# Load and Process OECD/UNESCO Data

```{r load-data}
# Open GERD data
gerd_data <- read_csv(file.path(data_dir, "r2", "btpsych_gerd_data.csv"))

# Open researchers data (skip first 4 rows)
researchers_data <- read_csv(
  file.path(data_dir, "r2", "btpsych_researchers_data.csv"), 
  skip = 4
)
```

# Calculate Averages for Year 2000 and Beyond

```{r calculate-averages}
# Process GERD data - calculate average for year 2000+
gerd_averages <- gerd_data |>
  filter(year >= 2000) |>
  group_by(iso3) |>
  summarise(
    avg_gerd = mean(gerd_value, na.rm = TRUE),
    .groups = "drop"
  )

# Process researchers data - calculate average for year 2000+
researchers_averages <- researchers_data |>
  filter(year >= 2000) |>
  group_by(iso3) |>
  summarise(
    avg_researchers = mean(researchers_value, na.rm = TRUE),
    .groups = "drop"
  )

# Combine GERD and researchers averages
combined_averages <- gerd_averages |>
  full_join(researchers_averages, by = "iso3")
```

# Load OpenAlex Data and Join

```{r join-data}
# Open OpenAlex processed data
openalex_data <- readRDS(file.path(data_dir, "openalex", "btpsych24a_data_bib_processed.Rds"))

# Join with combined averages using iso3 codes
final_data <- openalex_data |>
  left_join(combined_averages, by = "iso3")

# View summary of joined data
final_data |>
  select(iso3, avg_gerd, avg_researchers) |>
  summary()
```

# Data Overview

The analysis integrates three key datasets to examine the relationship between national science indicators and research output:

- **OECD GERD Data**: Gross domestic expenditure on research and development, representing national investment in R&D activities
- **UNESCO Researchers Data**: Number of research personnel, indicating human capital in scientific research
- **OpenAlex Bibliometric Data**: Publication and citation data from the bibliometric analysis

We calculated average values for both GERD spending and researcher counts for the period 2000 onwards to provide stable indicators of each country's research infrastructure and investment patterns.

```{r data-summary}
# Display country coverage
cat("Countries with GERD data:", nrow(gerd_averages), "\n")
cat("Countries with researchers data:", nrow(researchers_averages), "\n")
cat("Countries in final dataset:", nrow(final_data), "\n")

# Show distribution of key variables
final_data |>
  select(avg_gerd, avg_researchers) |>
  describe()
```

# Export Results

```{r export}
# Save the final joined dataset
saveRDS(
  final_data,
  file.path(data_dir, "processed", "btpsych_combined_with_indicators.Rds")
)

# Also save just the averages summary for reference
write_csv(
  combined_averages,
  file.path(data_dir, "processed", "country_averages_2000_plus.csv")
)
```

The final dataset combines bibliometric indicators with national science investment and human capital measures, enabling analysis of how research infrastructure relates to scientific productivity and impact in the field of psychology.